### 2021-02-13 action_letter.Rmd
Continued revision

---
title: "Action letter"
author: "Caspar J. van Lissa"
date: "2/11/2021"
output: html_document
---

# Reviewer #1:

## R1C1 The abstract is lacking in two regards:

### R1C1a The actual findings are not reported

### R1C1b it is not clear how the findings integrate te literature.

## R1C2 Key words are not provided.

Response: We apologize; the keywords were not retained when converting the manuscript to the required format for ADRR. We have reinstated them.

## R1C3 The title is about dysregulation but the first paragraph is all about regulation. The manuscript would be more effective if it were more consistent.

Response: We agree that greater consistency would improve the clarity of the writing, and have changed the manuscript to use the more general term "emotion regulation" throughout.

## R1C4 The first paragraph [promises] that the results will help unify theory and address limitations of research areas that the author deems in silos. This is a heavy burden to carry.

Response: We have interpreted this point as an indirect request that we further clarify the intended contribution of the present study. We will address that point below - but first we must clarify that it was not our intention to "promise that our results will help unify theory and address limitations of research areas that the author deams in silos". We have attempted to address this misunderstanding by clarifying that these are not our original claims; these are oft-repeated limitations of the existing literature, and our work is a response to these perceived limitations. We have included three references to support this point. The revision now reads:

> Despite the abundance of empirical work on this issue,
several authors have pointed out that the literature is somewhat fragmented,
because different (sub)disciplines have approached the issue in disparate ways,
in the absence of consistent terminology and unifying theory 
[@bussTheoriesEmotionalDevelopment2019; @bariolaChildAdolescentEmotion2011; @stifterEmotionRegulation2019].

Now, to the main point.
In the revision, we have attempted to clarify the intended contribution of the present study
as follows:

> There have been recent calls for the consolidation of empirical insights into overarching theory [e.g., @bussTheoriesEmotionalDevelopment2019]. 
According to formal methods for theory construction [@borsboomTheoryConstructionMethodologya; @haigExploratoryFactorAnalysis2005],
the first step in this endeavor is to identify the relevant *empirical phenomena*,
defined as stable and general features of the world [@bogenSavingPhenomena1988].
We propose that systematic reviews are a suitable method for taking this first step,
based on the assumption that phenomena currently considered to be most stable and general will be more frequently represented in the literature. [...] The present paper aims to map the empirical phenomena relevant to adolescent emotion regulation, using text mining systematic review.

## R1C5 The entire section on page 3: it is not clear why this is presented. It could be presented in a brief sentence and likely should because it does not really relate to what is reviewed. If it is kept, it should be edited so it relates to the actual review.
The paragraph on page 4 may be relevant, but it is not clear how it eventually relates to the study.

Response: These two paragraphs were intended to substantiate the relevance and scope of the present study.
Based on the Reviewer's comment, we now see that this was perhaps redundant with the opening paragraphs, which already provided rudimentary justification of the relevance and scope.
We have thus attempted to address the Reviewer's comment by cutting these paragraphs on pages 3 and 4,
merging the core arguments with the opening paragraphs.

The opening paragraphs now read:

> Adolescence is a life stage ranging from pubertal onset to adult-like independence [@steinbergAgeOpportunityLessons2014; @sawyerAgeAdolescence2018].
A key developmental challenge in this stage is acquiring mature emotion
regulation skills [@croneUnderstandingAdolescencePeriod2012;
@zimmermannEmotionRegulationEarly2014]. 
Adolescents experience biological, cognitive, and social changes that prompt new emotional
experiences and tax regulatory abilities [@steinbergAgeOpportunityLessons2014].
These coalescing changes temporarily restrict adolescents' capacity for emotion regulation 
[@zimmermannEmotionRegulationEarly2014],
resulting in more frequent and intense (negative) emotions [see @silkAdolescentsEmotionRegulation2003].
Although most adolescents eventually develop mature emotion regulation abilities,
for as many as one in five [@leeAdolescentMentalHealth2014],
adolescence marks the onset of
life course persistent emotion regulation-related psychopathology
[see also @aldaoEmotionregulationStrategiesPsychopathology2010;
@schaferEmotionRegulationStrategies2017].
Given the prevalence of emotion regulation difficulties in adolescence,
their implications for mental health
and downstream cost to society,
it is important to have a comprehensive understanding of the risk factors that render some adolescents more susceptible to emotional difficulties than others.

## R1C6 The existing theoretical landscape is presented in a way that is unclear how it really relates to the review.

Response: We understand that it is essential to clarify why we present this brief theoretical review before diving into the results of the present study.
In order to address this comment, we now explicitly state why a review of the existing theoretical landscape is relevant to the present systematic review. The revised manuscript reads:

> Although there is a paucity of theories *specific to* adolescent emotion regulation,
several relevant theories are commonly invoked in empirical work [see  @bussTheoriesEmotionalDevelopment2019].
In order to contextualize the results of our systematic review,
and assess to what extent the identified phenomena through text mining mirror or complement existing theory,
we provide a summary review of the existing theoretical landscape.

## R1C8 Figure 1 on page 5: it is not clear why it is there. It should come after the introduction. And, in fact, it should appear after the references and the text should simply have a placeholder for the Figure.  It should be referred to at the end of the introduction (before the current study section).

Response: As requested by the Reviewer, we have placed the Figure after the References section, retaining only a placeholder in-text. We also refer to the Figure at the end of the Introduction:

> As can be seen in Figure 1, 
there are notable parallels between correlates of emotion regulation identified in these
narrative reviews, and in the preceding theoretical literature.

## R1C9 The figure nicely points to outcomes, indicators, causes and protective factors. That is not clear from the introductory materials.

Response: In the previous manuscript, we had merged the two Figures from the theoretical Introduction with the two Figures from the Results section in order to facilitate comparisons between the two.
We see now that this caused some confusion.
In order to address this comment, we have separated the Figures from these two sections,
and we have added the following information about the classification of phenomena:

> In reading the theoretical literature,
we recorded relevant phenomena,
and whether these were discussed as putative causes, outcomes, indicators, or protective factors in relation to emotion regulation.
The left panel of Figure \@ref(fig:figbaseline) visually summarizes the phenomena so identified in the theoretical literature in a style similar to our subsequent systematic review,
in order to facilitate contextualizing our original findings.

## R1C10a The section on theory is not presented in a way that would have it appear as authoritative. It reads more like a hodgpode of theories.

Response: This is the state of the theoretical landscape: A hodgpodge. Others have pointed out that there is an absence of authoritative theories to guide research. It would be misleading to give readers the impression that there is a clear, authoritative theoretical framework.

## R1C10b The section would be stronger if it also gave readers a sense of which theories are deemed important/influential and worth using. There needs to be a sense that the important theories are being examined.

Response: Point out how often these theories are cited, perhaps?

## R1C11 The first paragraph on page 9 needs to be edited. It appears too haphazard. (Why use "Finally," as uo were not enumerating.) The problem, just as with the literature review, is that the studies appear presented randomly. One study did this, one did that, and so forth.

Response:

\ToDo

## R1C12 The fundamental problem with the "The present paper" section is that it really does not do what is needed. It is a plan of analysis. This section should note the research questions and hypotheses.

Response: We agree with the Reviewer that "The present paper" section was basically a plan of analysis. In the Revision, we have rewritten this section to summarize the study design, link it to the study aims, and state the intended outcomes. 
Note that this is an inductive (exploratory) study, and as such,
it does not test any hypotheses.
We now explicitly clarify this point in "The present paper".

> The present paper set out to map the phenomena relevant to adolescent emotion regulation
using a text mining systematic review (TMSR) of the literature.
This approach considers the frequency with which a phenomenon is covered in the literature to be indicative of its relevance,
and the frequency with which phenomena are investigated together within publications to be indicative of a relationship between these constructs.
The expected outcomes consist of term frequency metrics,
which can be rank ordered to identify the most relevant phenomena,
and term co-occurrence metrics.
These metrics can be jointly visualized as a network graph,
thereby "mapping" phenomena relevant to adolescent emotion regulation.
Note that this study is inductive or exploratory,
as opposed to deductive or confirmatory research, and as such does not test any hypotheses [see De Groot's work on the empirical cycle @degrootMethodologieGrondslagenVan1961][^Note that aside from induction, the term "abduction" is also used in the methodological literature. Injunction pertains to inferences based purely on statistical data, whereas abduction additionally involves the attribution of observations to explanatory principles.].

## R1C13 By the time I get to eh methods section, including the beginning of this section, I still see a focus on emotion regulation, not emotion dysregulation that appears in the title.

Response: We have addressed this comment as requested; see the response to R1C3.

## R1C13 As I read the materials on pages 11 and 12, I see a heavy reliance on the computerized analytical system to do the research and coding. I think it would be helpful to know what decisions are made by the authors and, if there are any, we should know about reliability checks and, if there are none, that still should be noted.

Response: This is a crucial comment with far-reaching implications.
The Reviewer asks what decisions are made by the authors.
Simply put: All decisions are made by us.
There are infinitely many ways to conduct exploratory analysis on a single data set,
and the present paper represents just one possible way.
The "computerized analytical system", as referred to by the Reviewer,
is nothing more than an explicit representation of our subjective decisions in analyzing the data,
represented as code.

The Reviewer asks for reliability checks, also known as a sensitivity analysis or multiverse analysis [@steegenIncreasingTransparencyMultiverse2016].
Such reliability checks are only defined, however, in the context of deductive (confirmatory) analysis,
where they are used to ensure that hypothesis tests / statistical inferences are robust to permutations of the data processing strategy.
As the present study is inductive (exploratory),
no hypotheses are tested nor inferences are drawn,
and the number of alternative possible analyses is infinite - thus reliability checks are undefined.

We do emphatically agree with the Reviewer that it is important to be clear about what decisions are made by us.

To address this comment, we have taken the following steps:

1) We have documented the methods in detail in the manuscript, and added online supplemental analyses for more complex analyses that we could not describe in sufficient detail.
2) We have made all analysis code and data publicly available in an open access repository.
This repository also contains the entire historical development of this study,
so readers can see that our analysis proceeded in a straightforward way. We mention this explicitly in the revised paper.
3) The revised Methods section now explicitly mentions that these are inductive (exploratory) analyses, and as such, depend on subjective decisions:

> As the purpose of inductive research is to inform hypothesis generation,
this approach affords the researcher with substantial creativity [@wagenmakersCreativityVerificationCyclePsychological2018].
This means that subjective decisions are made throughout the analysis process.
To ensure that all such decisions are properly documented,
we have made all code, data, and the historical record of this project publicly available in an online research repository at [https://github.com/cjvanlissa/veni_sysrev](https://github.com/cjvanlissa/veni_sysrev). 
We used the Workflow for Open Reproducible Code in Science [WORCS, @vanlissaWORCSWorkflowOpen2020]
to make the analyses strictly reproducible.
We encourage reuse of the analysis code,
and secondary analysis of the data.

## R1C14 At the bottom of page 12, it is clear that a decision is made in terms of retaining within a percentile. How do we know if that is the right decision?  There appear to be decisions like this made throughout. Given the reliance on the program, we need to know more.

Response: This is indeed a good example of a subjective decision.
In our response to the previous comment, we explain that we now explicitly acknowledge each such subjective decision in the revised manuscript.
In addition to the blanket statement we have added to the Methods section, 
we now also acknowledge the subjectivity of each decision like this one explicitly in the revised manuscript.

With regard to the Reviewer's question - "How do we know if that is the right decision?" - there is no straightforward answer.
We had thoroughly scoured the literature for established practice for pruning terms in text mining when preparing the initial submission, and did so again during revision,
but most published analyses appear to either retain a fixed number of terms, or drop terms that occur less than a fixed number of times.
Our selection method has the advantage of taking the distribution of word frequencies into account.
It is thus more flexible than a fixed cutoff, and still successful in pruning terms that occur only a few times in the corpus.
We now mention this explicitly in the paper:

> This is a subjective criterion,
but compared to the common practices of either retaining a fixed number of terms or pruning terms below a fixed frequency [see @benoitQuantedaPackageQuantitative2018],
it has the advantage of being responsive to the empirical distribution of term frequencies.

## R1C15 Page 13, there is again pruning. Justifications for the cutoff should be noted. There is a manual classification and we are left with no sense of its reliability/validity.

Please see our response to the previous comment.
Additionally, note that nearly all pruned terms occurred in only one document,
as can be seen in the plot below.
Even if the pruning criterion is subjective, it is hard to argue that we should interpret every word that occurred in the corpus.
Thus, we would argue, some pruning is necessary - and if a better method is available, we would be happy to use that instead.

```{r}
df_plot <- as.data.frame(term_freq)
df_plot <- df_plot[!df_plot$pruned, ]
ggplot(df_plot, aes(x = topterms, y = Freq)) + geom_bar(stat = "identity") + theme_bw() + labs(x = "Frequency of term", y = "Number of terms")
```

## R1C16 I really think the presentation would be more effective if it did not have results/discussions and then more results/discussions, and then a general discussion. We are not really dealing with different studies.

## R1C17 Page 20, the first full paragraph is a good example of text that is meant to show a contribution when it actually does not. Being the first or one of the first is not a contribution. What is more important is the actual finding. I realize that the focus here is on the method. But, being the first to use a method is not a contribution. What is important to highlight is the findings that come from the method, from the study.

## R1C18 Page 21, the political push for open science is not really a point that is relevant. If the authors wanted to review the actual papers, they could. They just need to identify the papers, download them, and search. Right? At any rate, I would avoid making political statements that, essentially, make the point that others should make their data available so that authors who use this method could use their method. Much of what is said here is not really reading well.

We understand that the Reviewer had interpreted this point as a political push for open science. It was not intended as such, and we have attempted to address this comment by rewriting this section. We have also attempted to clarify the important methodological implications of (lack of) open access for text mining analysis of scientific publications.
Availability of large datasets is crucial for machine learning approaches. As such, one of the major bottlenecks for more widespread / more in-depth application of text mining to scientific publication is the limited access to full text papers.

1) According to Archambault and colleagues (https://www.science-metrix.com/pdf/SM_EC_OA_Availability_2004-2011.pdf), only 36% of papers in the general social sciences are published in open access journals. In the present corpus, the proportion of papers published in directory of open access journals (DOAJ) was even lower, 10%. The remainder of publications is behind paywalls, thus requiring additional financial resources and manpower to access.
2) Even for papers that are available under open access licenses, there is no unified API for accessing the text. The reviewer's suggestion that authors can "just [...] identify the papers, download them, and search. Right?" is not scalable. Assuming that it would take only 10 minutes to find, download, and archive the full text, it would have taken 131.35 working days to assemble the corpus for the present study (6305 papers * 10 minutes)/(60 minutes * 8 hours per day). To manually curate a larger dataset, such as CORD-19 (publications about the coronavirus), would take 8333.33 days (400.000+ papers * 10 minutes)/(60 minutes * 8 hours per day).

## R1C19 in the end, it ends up looking like the authors are "hiding the ball". The real issue is what is actually being added to theory? What are the important findings here? What are the practical findings that emerge from this process that are worth making this type of analysis.  I just end up not convinced here.

Response: (hiding the ball = hiding evidence)

## R1C20 The conclusion note that the method charts "unknown territories". Sure that is not the case because the study relies on the published studies.

This is based on a misunderstanding about the level of analysis. We are not charting "unknown territories" with regard to predictors of emotion dysregulation, but with regard to which of those predictors are acknowledged as being relevant in the theoretical / review literature.

## R1C21 I really get a sense of over-selling. The authors should, again, focus on what is actually found that is useful about the topic. There is an example made of emerging themes and fringes, with the example given the little focus on fathers. That may be true, but there was not really a study needed to show that point, as it is well known that the field focuses on primary caretakers (mothers/women).  The conclusion ends with noting that others have noted a need for an overarching theory; but that is what is the apparent contribution being made here; the author does not really get to that and address it; there is no sense of a contribution being made to theory.

Our argument is precisely that the method used here identifies themes that