### 2021-04-06 manuscript.Rmd
finalize revisions

---
title             : "Mapping Phenomena Relevant to Adolescent Emotion Dysregulation: A Text-mining Systematic Review"
shorttitle        : "ADOLESCENT EMOTION REGULATION"

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"

abstract: |
  Adolesence is a developmentally sensititve period for emotion regulation
  with potentially lifelong implications for mental health and well-being.
  Although substantial empirical research has addressed this topic, the
  literature is fragmented across subdisciplines, and an overarching
  theoretical framework is lacking. The first step toward constructing a
  unifying framework is identifying relevant phenomena. This systematic
  review of 6305 papers used text mining to identify phenomena relevant to
  adolescents' emotion regulation. First, a baseline was established of
  relevant phenomena discussed in theory and recent narrative reviews. Then,
  article keywords and abstracts were analyzed using text mining,
  examining term frequency as an indicator of relevance and term
  co-occurrence as an indicator of association. The results reflected themes
  commonly featured in theory and narrative reviews, such as socialization
  and neurocognitive development, but also identified undertheorized themes,
  such as developmental disorders, physical health, external stressors,
  structural disadvantage, substance use, identity and moral development, and
  sexual development. The findings illustrate how text mining systematic
  reviews, a novel approach, may complement narrative reviews. Future
  theoretical work might integrate these undertheorized themes into
  an overarching framework, and empirical research might consider them as
  promising areas for future research, or as potential confounders in research
  on adolescents' emotion regulation.

keywords          : "emotion regulation, adolescence, systematic review, text mining, machine learning"
bibliography      : ["veni_sysrev.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man,noextraspace"
output            : papaja::apa6_pdf
knit              : worcs::cite_essential
header-includes:
  - \raggedbottom
---

```{r setup, include = FALSE}
library("papaja")
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(knitr)
library(kableExtra)
library(cowplot)
run_everything <- TRUE
notingraph <- list()
keyw <- function(x){
  x
}
cat_shapes <- c(Outcome = "square2", Indicator = "circle2", Cause = "quad2", Protective = "tri2")
legend_shapes <- c("circle2" = 21, "square2" = 22, "quad2" = 23, "tri2" = 24)[cat_shapes]
scale_shapes <- legend_shapes
names(scale_shapes) <- names(cat_shapes)[match(names(legend_shapes), cat_shapes)]
```
<!-- Reviewer suggestions:
Kristin Buss -->

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = FALSE, message = FALSE, warning = FALSE, results="hide")
```

Adolescence is a developmentally sensitive period for emotion regulation [@zimmermannEmotionRegulationEarly2014]. 
<!--Adolescents experience biological, cognitive, and social changes that prompt new emotional
experiences and tax regulatory abilities [@steinbergAgeOpportunityLessons2014].
These coalescing changes temporarily restrict adolescents' capacity for emotion regulation 
[@zimmermannEmotionRegulationEarly2014],
resulting in more frequent and intense (negative) emotions [see @silkAdolescentsEmotionRegulation2003].-->
Although most adolescents successfully develop mature emotion regulation abilities,
as many as one in five develop psychopathology [@leeAdolescentMentalHealth2014]
with roots in emotion regulation difficulties
[@aldaoEmotionregulationStrategiesPsychopathology2010].
Given the prevalence of emotion regulation difficulties in adolescence,
their implications for mental health<!--and social functioning [@braetEmotionRegulationChildren2014]; @reindlSocializationEmotionRegulation2016]-->
and downstream cost to society,
it is important to have a comprehensive overview of the phenomena associated with adolescents' emotion regulation.
<!--Of particular interest are potential risk factors that render some youth susceptible to emotion regulation difficulties.-->
Despite the abundance of publications in this field,
it is difficult to obtain a comprehensive understanding of emotion regulation in adolescence [@riedigerEmotionRegulationAdolescence2014]:
Different (sub)disciplines have approached the topic in disparate ways [@riedigerEmotionRegulationAdolescence2014],
without consistent terminology [@bariolaChildAdolescentEmotion2011], conceptual frameworks [@stifterEmotionRegulation2019], or an overarching theoretical framework
[@bussTheoriesEmotionalDevelopment2019].
This has prompted calls for the integration of knowledge across research areas [@riedigerEmotionRegulationAdolescence2014] and consolidation of empirical research into overarching theory [e.g., @bussTheoriesEmotionalDevelopment2019].
Formal methods for theory construction state that the first step in this endeavor is to identify well-established relevant *empirical phenomena*,
defined as stable and general features of the world [@borsboomTheoryConstructionMethodology2020].
The current study used a text mining systematic review to provide a comprehensive map of the empirical phenomena relevant to adolescent emotion regulation.

Currently, few instruments exist for the identification of relevant phenomena; instead, theorists typically rely on expert opinion [@borsboomTheoryConstructionMethodology2020].
This paper argues that systematic reviews are a suitable and relatively more objective method for phenomena detection.
Traditional narrative reviews are limited, however, by small convenience samples,
confirmation bias,
and an undue emphasis on positive results [@littellEvidencebasedBiasedQuality2008].
These limitations are overcome by a novel method:
the text mining systematic review (TMSR).
Relative to narrative reviews, text mining can cover arbitrarily greater corpora,
and gleans insights from the literature by means of a relatively more transparent, objective, and reproducible process.
<!--Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008],
the inherent "fairness" of text mining gives emerging themes a chance to come to the fore.-->
This approach assumes the frequency with which phenomena recur in the literature to be indicative of their relevance.
If it is additionally assumed that the frequency with which phenomena are studied together is indicative of the extent to which they are thought to be associated,
then the co-occurrence of phenomena within publications can be interpreted as a rudimentary nomological network: a mapping of relationships between theoretically relevant phenomena [see @alaviAligningTheoryMethodology2018].
Such a network can serve as a starting point for the formulation of a proto-theory,
which would additionally involve abductive reasoning: specifying putative mechanisms to explain associations between the phenomena [@borsboomTheoryConstructionMethodology2020].
Text mining systematic reviews constitute an inductive and exploratory approach:
Their aim is phenomena detection based on generalization from observations (data) without appeal to theory [@haigExploratoryFactorAnalysis2005].
By identifying and mapping the phenomena relevant to adolescent emotion regulation,
such an inductive approach lays the groundwork for future theory development.

<!--
stifterEmotionRegulation2019
* call for studies incorporating predictors of ER across multiple levels of analysis
* different conceptual frameworks

riedigerEmotionRegulationAdolescence2014
* difficulty in deriving a cohesive picture about emotion regulation in adolescence
* different conceptual frameworks

* three themes (neurocognitive development, socialization, interplay between internal and external factors) currently represent relatively independent bodies of research
* A more explicit integration of these various research perspectives will help us to arrive at a more integrated and interpretable picture of the development of emorion regulation during adolescence.

@bussTheoriesEmotionalDevelopment2019

* emotional development theory is still in its own state of development. 
* empirical evidence can contribute to theory development.

@bariolaChildAdolescentEmotion2011
* Further research apply reﬁned theoretical conceptualizations of ER.

Where sentient readers tend to structure their reading of the literature around established ideas [@littellEvidencebasedBiasedQuality2008],
This potentially reveals phenomena and associations missed by narrative reviewers,
thereby laying the groundwork for future empirical and theoretical work.

we set out to map these relevant empirical phenomena using a systematic review.

generate a proto-theoretical nomological network of the relevant empirical phenomena.
We discuss how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.
There have been calls for the  consolidation of these disparate insights into overarching theory would be an important step forward.
An important first step 
and intrinsic limitations of narrative reviews prohibit comprehensive coverage of the vast but diffuse literature .-->
<!--
Recent years have seen the publication of several excellent systematic reviews,
but these have all been narrative reviews.-->

 <!-- CJ To fill out gaps, following up on promising constructs and unify heretofore disparate regions of the lit -->
<!-- SH: this implies that the main contribution of the paper is predominantly methodological is that the case? Is there some primary thesis/argument that will be advanced? It is really important to take a top-down approach to that messaging, beginning here in the first paragraph-->
<!-- Ergens anders: Similarly, although
there are several *relevant* theories, there is no specific theoretical
framework to guide research on adolescents' emotion regulation. Although theory
development is beyond the scope of this paper, we set out to generate a
proto-theoretical nomological network of the potential risk factors,
manifestations, and outcomes of emotion regulation in adolescence. We discuss
how this nomological network relates to existing relevant theory, and conclude
with recommendations for future theory-generating efforts.-->
<!-- At the same time, the staggered development of motivational-emotional and regulatory brain circuits gives rise
to a maturity gap [@croneUnderstandingAdolescencePeriod2012]. 
This leads youngsters to pursue new experiences in life and love,
without being fully prepared to cope with the emotional outcomes. -->

<!--The literature thus paints a picture of
adolescence as a chrysalis for emotional development: Children enter this stage
with emotion regulation skills adapted to the challenges of childhood. During
adolescence, emotional systems are rearranged substantially.-->

<!--
* intense negative emotions in daily life (Silk, Steinberg, & Morris, 2003) 
* unstable peer or romantic relationships (Furman & Collins, 2009)
* decrease in perceived support from parents (Furman & Buhrmester, 1992).
* early adolescence is characterized by a higher rate of conflicts with parents (Laursen, Coy, & Collins, 1998)
* higher variability of negative emotions compared to late adolescence (Larson, Moneta, Richards, & Wilson, 2002).
* relative to other developmental stages, adolescence involves a particularly large number of transitions, novel situations, and new stimuli, physical, cognitive, emotional, and social development—particularly in early and mid-adolescence (Seiffge-Krenke, 2000)
-->

## Theoretical Approaches to Emotion Regulation

```{r baselinecode, message=FALSE, echo = FALSE, warning = FALSE, error=FALSE, results='hide'}
library(data.table)
library(stringr)
library(udpipe)
library(igraph)
library(wordcloud)
library(Matrix)
library(ggplot2)
library(yaml)
source("word_functions.R")
source("circle2.R")
dict <- read_yaml("yaml_dict.txt")

  bl_theory <- readLines("baseline_theory.txt")
  bl_review <- readLines("baseline_reviews.txt")
  df <- data.frame(word = trimws(tolower(bl_theory)), source = "theory")
  df <- rbind(df, data.frame(word = trimws(tolower(bl_review)), source = "reviews"))
  # Categorize words
  res_cat <- cat_words(df$word, dict, handle_dups = "all")
  df$cat <- res_cat$words
  df <- df[!duplicated(df$cat), ]
  if(!is.null(res_cat[["unmatched"]])){
    df <- df[!df$cat %in% names(res_cat$unmatched), ]
  }

  baseline_cat <- df
  baseline <- c("dysregulation", unique(df$cat[df$source == "theory"]))

  saveRDS(baseline, "baseline.RData")
  saveRDS(baseline_cat, "baseline_cat.RData")

in_theory <- df$word[df$source == "theory"]
in_rev <- df$word[!df$source == "theory"]
in_rev[!in_rev %in% in_theory]

df_plot <- data.frame(term1 = df$cat[df$source == "theory"], term2 = "dysregulation")
df_plot <- df_plot[!df_plot$term1 == df_plot$term2, ]
edg <- df_plot

edg$width = 1

vert <- data.frame(name = unique(c(df_plot$term1, df_plot$term2)), label = pretty_words(unique(c(df_plot$term1, df_plot$term2))), size = 1)
vert <- vert[vert$name %in% unique(c(edg$term1, edg$term2)), ]

categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
if(any(!vert$name %in% categ$name)){
  write.table(vert$name[!vert$name %in% categ$name], "clipboard", sep = "\n", row.names = FALSE, col.names= FALSE)
  stop("Please re-categorize missing vertices.")
} 
vert$Category <- categ$category[match(vert$name, categ$name)]

cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "forestgreen")
cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "olivedrab2")
vert$color <- cat_cols[vert$Category]
vert$shape <- cat_shapes[vert$Category]
vert$frame.color <- cat_cols[vert$Category]

vert$size <- scales::rescale(log(vert$size), c(4, 12))
g1 <- graph_from_data_frame(edg, vertices = vert,
                           directed = FALSE)

# Plot 2 ------------------------------------------------------------------

df_plot <- data.frame(term1 = df$cat, term2 = "dysregulation", source = df$source)

edg <- df_plot

edg$width = 1

vert <- data.frame(name = unique(c(df_plot$term1, df_plot$term2)), label = pretty_words(unique(c(df_plot$term1, df_plot$term2))), size = 1)
vert <- vert[vert$name %in% unique(c(edg$term1, edg$term2)), ]

categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
if(any(!vert$name %in% categ$name)){
  write.table(vert$name[!vert$name %in% categ$name], "clipboard", sep = "\n", row.names = FALSE, col.names= FALSE)
  stop("Please re-categorize missing vertices.")
} 
vert$Category <- categ$category[match(vert$name, categ$name)]
vert$faded <- vert$name %in% df$cat[df$source == "theory"]
cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "olivedrab2")
vert$color <- cat_cols[vert$Category]
vert$shape <- cat_shapes[vert$Category]
vert$frame.color <- cat_cols[vert$Category]

vert$color[vert$faded] <- "#FFFFFF"

vert$size <- scales::rescale(log(vert$size), c(4, 12))
g2 <- graph_from_data_frame(edg, vertices = vert,
                           directed = FALSE)

set.seed(4) #4 #2 #3
l1 <- l <- layout_with_fr(g1)
set.seed(3)
l2 <- layout_with_fr(g2)

p <- quote({
  # Set margins to 0
  par(mar=c(0.5,0.5,0.5,0.5),
      mfrow=c(1,2))
  plot(g1, edge.curved = 0, layout=l1,
       vertex.label.family = "sans",
       vertex.label.cex = 0.7,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "a")
  legend(x=-1.2, y=1.1, names(cat_cols), pch=legend_shapes, col=cat_cols, pt.bg=cat_cols, pt.cex=1.4, cex=.7, bty="n", ncol=1)
  plot(g2, edge.curved = 0, layout=l2,
       vertex.label.family = "sans",
       vertex.label.cex = 0.7,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "b")
})
saveRDS(g1, "bl_g1.RData")
saveRDS(g2, "bl_g2.RData")
saveRDS(l1, "bl_l1.RData")
saveRDS(l2, "bl_l2.RData")
# Save files

png("baseline_network.png", height = 100, width = 200, units = "mm", res= 300)
eval(p)
dev.off()

svg("baseline_network.svg", width = 14)
eval(p)
dev.off()
```

```{r figbaseline, warning=FALSE, eval = TRUE, results='asis', out.width = "100%", fig.cap="Phenomena relevant to adolescents' emotion regulation according to theory (a) and narrative reviews (b; transparent nodes indicate constructs also present in the theory)."}
knitr::include_graphics("baseline_network.png") 
```
```{r allnetworks, warning=FALSE, eval = FALSE, results='asis', out.width = "100%", fig.cap="Correlates of emotion regulation difficulties according to theory (a), narrative reviews (b), and term co-occurrence in author keywords (c) and abstracts (d). Size of lines and nodes represents frequency. Dashed lines represent links not involving dysregulation. Transparent nodes indicate constructs also represented in the theory."}

knitr::include_graphics("networks.png") 
```

Although it has been argued that there is a paucity of theories *specific to* adolescent emotion regulation [see @bussTheoriesEmotionalDevelopment2019],
several *relevant* theories are commonly invoked in empirical work.
A brief review of the existing theoretical landscape is provided here
to contextualize the present systematic review
and assess to what extent phenomena identified through text mining mirror or complement existing theory.
Others have published detailed reviews of theories of emotional development [@bussTheoriesEmotionalDevelopment2019] and of empirical research on adolescent emotion regulation [@riedigerEmotionRegulationAdolescence2014].
The left panel of Figure \@ref(fig:figbaseline) visually summarizes the phenomena discussed in the theoretical literature reviewed below,
along with whether these were discussed as putative causes, outcomes, indicators, or protective factors in relation to emotion regulation.
This establishes a baseline for the original contributions of the text minning analysis.
<!-- 
Our text mining analysis of the phenomena empirical literature is best understood against a backdrop of -->

Two of the most general theories invoked to frame developmental research are the bioecological model [@bronfenbrennerBioecologicalModelHuman2007], and the transactional model [@sameroffUnifiedTheoryDevelopment2010].
The bioecological model describes how the `r keyw("environment")`
shapes individual development.
At the individual level, every person is imbued with `r keyw("biological")` predispositions,
and develops over time in interaction with contextual influences.
The most immediate source of contextual influences is the microsystem,
composed of primary caregivers and other close ties.
Other environmental influences originate in the macrosystem - including `r keyw("political")` and `r keyw("economic")` influences - and the exosystem,
consisting of `r keyw("cultural")` norms and `r keyw("values")`.
The transactional model [@sameroffUnifiedTheoryDevelopment2010] is compatible with the bioecological model,
but places a stronger emphasis on development as a product of reciprocal influences between child and environment.
The transactional model distinguishes between proximal influences, which roughly correspond to the microsystem in the bioecological model, and distal influences, which derive from
structural and environmental factors indirectly shaping development, like `r keyw("socio-economic status")`,
`r keyw("schools")`, and the `r keyw("community")` (macro- and exosystem).
With increasing age, these distal influences gain ground on proximal influences.
These two theories have a broader scope than most.
This means that they can be invoked to contextualize any developmental study,
but lack specificity - a shortcoming that curtails a theory's utility in generating hypotheses.
Domain-specific theories, by comparison, offer greater specificity.

One such domain-specific theory is Hall's notion of "storm and stress" in adolescence [see
@arnettAdolescentStormStress1999].
It describes how `r keyw("hormonal")` changes diminish `r keyw("self-control")` and increase `r keyw("reactivity")`, which in turn leads to emotion
regulation difficulties, increased `r keyw("conflict with parents")`, and `r keyw("risky behavior")`.
Contemporary theories recast this notion of diminished self-control and increased emotional reactivity as a normative change that facilitates emotional maturation at the risk of emotional
disturbance [@arnettAdolescentStormStress1999].
A potential limitation of conceptualizing emotion regulation difficulties as normative is that this underemphasizes heterogeneity in developmental trajectories [cf. @croneUnderstandingAdolescencePeriod2012].
<!--SH I am still not sure what the empirical contribution is - can you be more specific with it in this sentence? -->
<!--Several recent publications have undertaken the
monumental task of providing a comprehensive overview of theories of emotional
development [e.g., @bussTheoriesEmotionalDevelopment2019;
@holodynskiDevelopmentEmotionsEmotion2006;
@riedigerEmotionRegulationAdolescence2014], and of the empirical work
pertaining to emotions in adolescence
[@coe-odessEmergentEmotionsAdolescence2019].  -->

One theory of normative emotional development focuses on developmental influences to a greater extent [@sroufeEmotionalDevelopmentOrganization1995].
It posits that, as children grow older,
their increasing `r keyw("self-regulatory")` abilities
drive a transition from external emotion regulation by `r keyw("primary caregivers")`
toward autonomous emotion regulation.
This theory emphasizes `r keyw("social")` and `r keyw("cognitive influences")` as 
drivers of emotion regulation development.
Social influences mainly occur
through `r keyw("parental co-regulation")`, `r keyw("parenting behaviors")`, and parent-child
`r keyw("attachment")`.
Cognitive influences occur through the development of the `r keyw("central
nervous system")` (CNS), `r keyw("cognition")`, and `r keyw("self-regulation")`.
<!--In terms of level of analysis, Sroufe's and Hall's theories corresponds approximately to the individual- and microsystem level of the ecological model [@bronfenbrennerBioecologicalModelHuman2007].-->
This theory's relevance for adolescent research is somewhat diminished by its focus on early childhood.
However, research indicates that there is some continuity in developmental influences throughout childhood and adolescence; for instance, mothers' and fathers' unique roles in emotional development are similar in childhood and adolescence [see @vanlissaMothersFathersQuantitative2020; @vanlissaRoleFathersMothers2019].
The emphasis on socialization and `r keyw("neurocognitive development")`,
which is evident from both theories,
is the focus of several theories with a narrower scope.

One widely cited theory of socialization is the tripartite model [@morrisRoleFamilyContext2007].
It describes three pathways through which `r keyw("parents")` shape emotion regulation: `r keyw("modeling")`, `r keyw("parenting practices")`, and the
emotional `r keyw("family climate")`,
which subsumes attachment and `r keyw("marital relationship")` quality.
This theory acknowledges unique contributions by `r keyw("mothers")` and `r keyw("fathers")`, and the importance of siblings.
Others have adapted the tripartite model to describe peer socialization [@reindlSocializationEmotionRegulation2016], thus illustrating its wider usefulness.
A more abstract take on socialization is found in the internalization model of emotional development [@holodynskiDevelopmentEmotionsEmotion2006]. 
This theory describes the socialization of the role of emotion in `r keyw("communication")`, and the `r keyw("cultural")` symbolic function of emotion.

In research on neurocognitive aspects of adolescents' emotion regulation,
`r keyw("polyvagal theory")` has been influential [@porgesOrientingDefensiveWorld1995].
This theory examines `r keyw("emotional experience")` and -regulation in relation to autonomous nervous system functioning, respiratory sinus arrhythmia, and the stress response.
Although polyvagal theory is relevant for development,
it does not explicitly address it.
Among developmental neurocognitive theories,
there is some consensus that the developmental asymmetry between `r keyw("motivational")`-emotional and inhibitory `r keyw("brain")` circuits gives rise to a "maturity gap" in middle adolescence [@croneUnderstandingAdolescencePeriod2012].
According to the model of social-affective engagement and `r keyw("goal flexibility")` [@croneUnderstandingAdolescencePeriod2012],
adolescents' cognitive engagement is dynamically responsive to
social and motivational `r keyw("goal salience")`.
This flexibility <!--, on the one hand,-->
prepares adolescents to <!--effectively engage cognitive systems in novel
challenging situations in a way that facilitates--> develop mature regulatory
abilities, but also places them at risk of greater impulsivity - for example, in
pursuit of `r keyw("peer approval")`.
To a greater extent than related writings [cf. @craccoEmotionRegulationChildhood2017], this model focuses on adolescents diverging destinies;
why some youngsters flourish while others languish [@croneUnderstandingAdolescencePeriod2012].
This is important for understanding individual differences in development.
Although the relevance of this theory is enhanced by its focus on adolescence,
it only tangentially addresses emotion regulation.
Furthermore, whereas this theory addresses cognitive factors and peers as drivers of development, it underemphasizes other relevant phenomena, such as parenting [cf. @morrisRoleFamilyContext2007].
Micro-scale theories like these are useful in explaining specific phenomena in detail,
but do not provide a comprehensive understanding of adolescent's emotion regulation.
To this end, it is important to consider them in the context of a larger framework of relevant phenomena.

<!--Like Sroufe, Holodynski describes an age graded transition from
interpersonal to intrapersonal emotion regulation.
In doing so, Holodynski applies Vygotsky's theory of development to the domain of
emotion, and presents a detailed integrated model of emotional experience and
regulation.-->
Theories of emotion regulation offer further insight into
intra-individual drivers of emotion regulation development.
The influential process model describes emotion regulation throughout the progression from eliciting cue to ultimate response [@grossHandbookEmotionRegulation2013].
This model posits that individuals use `r keyw("strategies")` to modulate this process at different stages,
consciously or otherwise. 
Maladaptive emotion regulation strategies tend to result in more negative emotions,
diminished well-being, and greater strain in interpersonal relationships
[ @bellRelationshipsInputsOutputs2000].
However, comparative studies indicate that the adaptive versus maladaptive psychosocial consequences of specific strategies appear to be partly contingent on the `r keyw("cultural context")` [see
@bariolaChildAdolescentEmotion2011].
Similar to the process model, the "social
information processing" theory also describes the role of cognitive
processes and strategies in emotion regulation
[@lemeriseIntegratedModelEmotion2000].
One shortcoming of these theories for understanding adolescent emotion regulation 
is the lack of a developmental component.

<!-- RK: What do you take from these broad perspectives for your RQ?-->
<!--SH I am not sure what contribution this paragraph ultimately makes - you should consider condensing it substantially to just note that these theories have something to say about emotional development but are not adolescent-specific-->
<!--Theories like this are particularly relevant to a fine-grained
understanding of the process of emotion regulation, but their developmental
relevance has unfortunately not yet been considered (REF Buss). -->
<!--Parents are widely considered to be the primary proximal influence driving
emotion regulation development.
Yet several limitations remain.
@coe-odessEmergentEmotionsAdolescence2019;
instead takes an inductive approach to map the correlates of adolescents'
emotion regulation based on all available literature.

The present study instead takes an inductive approach to map the correlates of
adolescents' emotion regulation based on all available literature. From this
mapping, we proceed to classify correlates as potential risk factors,
manifestations, and outcomes of adolescent emotion regulation. The results
provide a conceptual overview of the existing literature, can help identify
blind spots in existing theory, and might inspire new hypotheses to guide
future deductive research.

Several limitations emerge from prior efforts to provide an encompassing
framework of this literature

These efforts at unification have, thusfar, been conducted in a top-down,
theory-driven manner.

which motivates most youngsters to seek out challenges in life and love, but
renders some of them vulnerable to emotion dysregulation.  -->

<!--
Note that the left panel of Figure \@ref(fig:tmnetworks) visually summarizes the correlates of emotion regulation difficulties (labeled: *Dysregulation*) according to our theoretical review.
The left panel of Figure \@ref(fig:tmnetworks) summarizes key constructs covered by the theories reviewed above. 
In the subsequent text mining analysis, we will distinguish the topics covered in this theoretical review visually to highlight unique contributions of our approach.
This allows us to see in what respects the text mining analysis reflects existing theory,
and in what respects it complements it.-->
<!--### Shortcomings of existing theory-->

Despite the abundance of theory relevant to emotion regulation in adolescence,
the literature has several limitations.
First, few theories have explicitly addressed adolescence.
As this life stage differs qualitatively from both childhood and adulthood,
it is questionable whether theories focused on different age groups can be generalized to adolescents [@bariolaChildAdolescentEmotion2011].
<!--Relatedly, many relevant theories lack an explicit developmental component, which renders them and thus lack a well-substantiated
understanding of developmental processes[see
@bussTheoriesEmotionalDevelopment2019;
@croneUnderstandingAdolescencePeriod2012]. -->
Furthermore, few theories have comprehensively addressed important predictors of development in this life stage,
and none directly guide contemporary research in the field [@bussTheoriesEmotionalDevelopment2019].
Finally, existing theories vary widely in scope:
Some are broad and non-specific; others describe a specific phenomenon in detail,
but lack a broader perspective.
Broad theories can be used to frame a wide range of research,
and specific theories are more suitable for deductive hypothesis generation.
To combine the strength of both, it would be beneficial to bridge these levels of analysis.
There is a need for integrative theory formation,
which could provide a unified framework to guide future empirical work.<!--[see @riedigerEmotionRegulationAdolescence2014].
Inductive analysis of the empirical literature may constitute a first step to this end.-->
<!--Such a theory should focus on -->
<!--
Barriola: What is most
* evident from the reviewed studies is that the samples
investigated have been mainly non-clinical, within the
early childhood period of development and have examined
predominantly maternal socialization factors.
* the vast
majority have been restricted to toddlers and young chil-
dren and have only examined the in?uence of maternal,
and not paternal, expression.
Most of the literature has addressed emotion regulation in adult or infant
samples. A smaller, but substantial, portion of the literature has examined
emotion regulation in adolescents.
-->

<!-- Next paper: * Existing theory is not very specific, in the sense that it
can give rise to concrete, testable predictions. "either is not guided by
hypotheses of an explicit theory or cites a theory but does not explicitly test
the theory." (Buss) (Smaldino)-->
<!--Unfortunately, there is little theory about which risk factors and
environmental hazards render adolescents susceptible to emotional
difficulties11-->

<!--Initially, much of the research on emotion regulation focused on either
adulthood or early childhood. In recent years, however, a substantial body of
research on emotion regulation in adolescence has accrued.-->

<!--What are the predictors of emotion regulation in adolescence?-->

<!--Emotion regulation can be defined as the ability to modulate emotional
experience and expression through automatic or deliberate processes. -->

<!--

## Why focus on theory?

A focus on theory is becoming increasingly important, given the concern about
the replicability of social scientific findings [@KLEIN].

Several authors of comprehensive reviews have noted that developmental theories
of emotion are still in their infancy [@bussTheoriesEmotionalDevelopment2019].
Several specific limitations of the theoretical landscapes precipitate the
present study. First, although there are many *relevant* theories, few are
focused specifically on emotion regulation. Second, the stage of adolescence is
rarely addressed, although empirical research has indicated that this is a
crucial period of emotion regulation development.
"emotion theory has largely focused on adults, and emotional development theory
has largely focused on infancy and early childhood (see Chap. 24 on adolescent
emotional development)." Buss et al 2019:20

## Why inductive -->

## Prior Narrative Reviews

<!--Whereas theory provides a top-down deductive understanding of adolescents' emotion regulation,-->
Literature reviews can shed light on undertheorized relevant phenomena by
synthesizing inductive insights emerging from the empirical literature.
This is especially important given the noted absence of overarching theory [see @bussTheoriesEmotionalDevelopment2019].
One recent narrative review, in exploring which factors contribute to adolescents' emotion regulation development, clearly reflected the emphasis on neurocognitive factors and socialization that is evident from the theoretical literature [@riedigerEmotionRegulationAdolescence2014].
Other recent narrative reviews complement the theoretical literature to a greater extent by addressing relevant but undertheorized phenomena [@coe-odessEmergentEmotionsAdolescence2019; @bariolaChildAdolescentEmotion2011].
The right panel of Figure \@ref(fig:figbaseline) illustrates in what respects these narrative reviews reflect existing theory,
and in what respects they complement it.
<!--
Note that the left panel of Figure \@ref(fig:tmnetworks) visually summarizes the correlates of emotion regulation difficulties (labeled: *Dysregulation*) according to our theoretical review.
The left panel of Figure \@ref(fig:tmnetworks) summarizes key constructs covered by the theories reviewed above. 
In the subsequent text mining analysis, we will distinguish the topics covered in this theoretical review visually to highlight unique contributions of our approach.
This allows us to see in what respects the text mining analysis reflects existing theory,
and in what respects it complements it.-->

One seminal review discussed relevant phenomena at different levels of analysis [@bariolaChildAdolescentEmotion2011]. 
At the individual level, such relevant phenomena include temperament and biological factors, like neurocognitive development and genes.
At the level of proximal influences, the authors discuss socialization and modeling by parents, teachers, and peers.
Finally, at the level of distal influences, culture and the media are discussed.
This review also called for research on emotion regulation beyond early childhood and into adolescence, and on fathers' unique contributions -
a topic receiving increasing interest [see @pleckPaternalInvolvementRevised2004].
In line with these recommendations, several recent publications have investigated mothers' and fathers' unique roles in emotion regulation socialization from childhood to adolescence [see @vanlissaMothersFathersQuantitative2020] and throughout adolescence [see @vanlissaRoleFathersMothers2019].

As the aforementioned review is now a decade old,
it is worth considering a more recent review for additional insight [@coe-odessEmergentEmotionsAdolescence2019]. 
This review complements offers a nuanced discussion of several additional topics.
At the individual level, this includes the implications of physiological changes,
including neurocognitive development and pubertal maturation.
Pubertal development also precipitates sexual and romantic behavior,
and the intensification of both biological sex differences and gender stereotyped behavior.
This, in turn, likely modulates proximal influences through peers.
<!--With regard to the process of emotion regulation,
Coe-Odess and colleagues discuss the importance
of strategies, and point out that, in addition to negative emotionality,
positive emotionality also peaks in adolescence. -->
The review further describes how individual hormonal changes intensify the stress response,
which relates to adolescents' greater susceptibility to emotion regulation difficulties.
Finally, going beyond the implications of cognitive development
discussed in other publications, the authors discuss how cognitive development and increased capacity
for abstract thought relate to identity formation and increased emotional
understanding.
These themes relate to two key challenges in adolescence, namely identity [@meeusStudyAdolescentIdentity2011] and empathy development [@vanlissaLongitudinalInterplayAffective2014].
At the level of proximal influences, the review expands on the role of conflict with parents, which peaks in adolescence as youth become increasingly individuated.
This is relevant because parent-adolescent conflict has been shown to impact both day-to-day mood swings and dispositional emotion regulation [see @vanlissaCostEmpathyParentadolescent2017].
The review further describes how adolescents become increasingly oriented toward peers, which
increases their sensitivity to social status and norms,
and prompts concomitant increases in peer pressure and risk taking.
The review devotes limited attention to distal influences.

There are notable parallels between phenomena relevant to adolescents' emotion regulation as identified in these
narrative reviews, and in the preceding theoretical literature, as can be seen in Figure \@ref(fig:figbaseline).
Nevertheless,
these literature reviews also touch upon issues that have received little
attention in theoretical work.
<!--These include specific individual differences at the biological and psychological level, including genetic predisposition, hormones, pubertal onset, gender, sexuality, temperament, identity, empathy, and stress.
Further introduced are proximal influences, including the role of media, norms, and social status.-->
This illustrates the general principle that reviews of the empirical literature can contribute inductive insight into relevant phenomena underrepresented in theory.
An important shortcoming is that all reviews in this field have been unstructured narrative reviews; a method that is often limited in scope and susceptible to bias [@littellEvidencebasedBiasedQuality2008].

<!--Where sentient readers tend to structure their reading of the literature around established ideas,
text mining has the potential to reveal phenomena and associations missed by narrative reviewers,
thereby laying the groundwork for future empirical and theoretical work.-->

<!-- NIET TEVEEL OVER THEORIE PRATEN; IK KIJK NIET NAAR THEORIE. IK KIJK NAAR FACTOREN DIE IN VERBAND GEBRACHT WORDEN MET EMOTIEREGULATIE. -->
<!--The life stage of adolescence differs qualitatively from both childhood and adulthood. It is therefore not sufficient to extend -->

<!--
Influences on emotion regulation (Riediger Klipker)
* Neurophysiological development 
* Familial context

Despite widespread interest in emotion regulation in adolescence, there is a lack of explicit theories about which risk factors and environmental hazards render adolescents susceptible to emotional difficulties. Some theories can be considered relevant, but these theories are rarely explicit enough to generate specific predictions. 
As many as one in five adolescents experiences severe emotional problems. A key question is therefore which factors render adolescents at-risk for difficulties in emotion regulation.

Unfortunately, relevant theory on adolescent emotion regulation development is limited27–29. The field is therefore at an impasse: We know that some adolescents are more susceptible to emotion dysregulation than others, but lack tools and theory to identify important predictors of individual development30.

inductive approach, thereby facilitating a more complete understanding of developmental processes, and nourishing theory formation

My work addresses this lacuna by taking an inductive approach to theory formation; paying special attention to between-person differences. This paves the way for a new wave of person-centered research

Key questions:

Why focus on emotion regulation specifically, instead of other aspects of emotional development?

* Biological determinants
* Social determinants

* Existing developmental theories do not focus on adolescence (rather, on adults, or development in infancy and childhood). However, many interesting emotion-related things happen in adolescence.

Important:
Buss: theories of emotion and emotional development
Cracco: Evidence for a maladaptive shift in emotion regulation in adolescence
Zimmermann & Iwanski: Dip in adolescence

Sroufe 1995: Development of emotion regulation until preschool age
"the key progression in emotional regulation is from caregiver-orchestrated
regulation to dyadic regulation to self-regulation."
* attachment - dyadic regulation of emotion (Sroufe 1996)
* Caregiver-guided self-regulation 
* Autonomous self-regulation
***** HOWEVER: In adolescence, regulatory brain regions still developing!!!! So Sroufe stops too early

Polyvagal theory Porges 1995 -->

<!-- SH: ultimately I think that the introduction is too long - considering that the introduction, itself, is a literature review to some extent, I don't come away with a clear delineation between the literature review up this point and the new work you are going to contribute. I think you need to review the Introduction with an eye on focusing the discussion on the problems/gaps that exist and how the lit review method usually used are insufficient for addressing those issues-->

# The Current Study

The empirical literature on adolescent emotion regulation is somewhat fragmented accross subdisciplines, and lacks an overarching theoretical framework.
The current study seeks to integrate this body of work by means of a comprehensive text mining systematic review (TMSR).
This is an inductive study, and as such,
does not test hypotheses [@wagenmakersCreativityVerificationCyclePsychological2018].
The aims of the study are to identify the most relevant phenomena in the body of literature,
based on the frequency with which phenomena are reported,
and to map potential relationships between phenomena, based on their co-occurrence within studies.
The resulting map of relevant phenomena helps integrate the empirical literature.
Moreover, as identifying relevant empirical phenomena is the first step in theory construction methodology,
this study lays the groundwork for future theory development.
<!--The distinction between inductive and deductive research has recently been recast as a "creativity-verification" cycle [@wagenmakersCreativityVerificationCyclePsychological2018].-->

# Methods

This text mining systematic review considered the frequency with which a phenomenon is covered in the literature, or *term frequency*, to be indicative of its relevance.
Term frequency can be rank ordered to identify the most relevant phenomena.
Furthermore, the frequency with which phenomena are investigated together within publications, or *term co-occurrence*, is considered to be indicative of a putative relationship between them.
These two metrics can be jointly visualized as a network graph,
thereby "mapping" phenomena relevant to adolescent emotion regulation.

Inductive research affords the researcher with substantial creativity [@wagenmakersCreativityVerificationCyclePsychological2018].
This means that subjective decisions are made throughout the analysis process.
To ensure that all such decisions are properly documented,
all code, data, and the historical record of this project are available in a public research repository at [https://github.com/cjvanlissa/veni_sysrev](https://github.com/cjvanlissa/veni_sysrev). 
The Workflow for Open Reproducible Code in Science was used to make all analyses reproducible [WORCS, @vanlissaWORCSWorkflowOpen2020].
Reuse of the analysis code and secondary analysis of the data are encouraged.

## Search strategy

The search was conducted in Web of Science,
the most comprehensive database accessible to the lead author with permissions to export keywords and abstracts.
The search strategy was based on
procedures described by Staaks [@staaksSystematicReviewSearch2019].
First, a reference set of 29 articles was compiled.
Then, a search string was constructed to retrieve the articles in this set.
The search string consisted of synonyms of emotion regulation and adolescence.
It returned 6653 results, including 25 records in the reference set.
To match all 29 reference set items required adding the terms `"emotio* socialization" OR "emotio* processes"` as synonyms for emotion regulation.
Doing so resulted in 191 more hits, most of which did not meet the inclusion criteria explained below.
These terms thus appeared to be overly inclusive, and the original search string was used.
```{r readrecs}
filenam <- "recs_6653.RData"
if(!file.exists(filenam)){
  library(bibliometrix)
  f <- list.files("./recs/18-3-2020/", full.names = TRUE)
  f <- readFiles(f)
  recs <- convert2df(f)
  saveRDS(recs, filenam)
} else {
  recs <- readRDS(filenam)
}
```

```{r removeduplicates}
filenam <- "unique_recs.RData"
if(!file.exists(filenam)){
  library(revtools)
  recs$doi <- recs$DI
  recs$doi[is.na(recs$doi)] <- paste0("fake", 1:sum(is.na(recs$doi)))
  if(any(duplicated(recs$doi))){
    recs$doi[duplicated(recs$doi)|duplicated(recs$doi, fromLast = T)]
    dup_doi <- duplicated(recs$doi)
    recs <- recs[!dup_doi, ]
  }
  recs$title <- recs$TI
  # locate and extract unique references
  potential_dups <- find_duplicates(recs, match_variable = "title", match_function = "stringdist", threshold = 5, to_lower = TRUE, remove_punctuation = TRUE)
  unique_recs <- extract_unique_references(recs, matches = potential_dups)
  saveRDS(unique_recs, filenam)
  write.csv(data.frame(dup_doi = sum(dup_doi), dup_title = (nrow(recs)-nrow(unique_recs))-sum(dup_doi)), "dups.csv", row.names = FALSE)
} else {
  unique_recs <- readRDS(filenam)
  dups <- read.csv("dups.csv")
}
unique_recs$id_num <- 1:nrow(unique_recs)

if(FALSE){
  # df_screen <- unique_recs[, c("id_num", "TI", "AB")]
  # names(df_screen) <- c("id_num", "title", "abstract")
  # write.csv(df_screen, file = "asreview.csv", row.names = FALSE, fileEncoding = "UTF-8")
  df_screen <- unique_recs[, c("id_num", "TI", "AU", "SO", "SN", "VL", "IS", "BP", "PY", "PU", "DI", "AB")]
  names(df_screen) <- c("key", "title", "authors", "journal", "issn", "volume", "issue", "pages", "year", "publisher", "url", "abstract")
  write.csv(df_screen, file = "rayyan.csv", row.names = FALSE)
}
rayyan_res <- read.csv("rayyan_exports/articles.csv")
rayyan_dups <- nrow(unique_recs) - nrow(rayyan_res)

```

## Screening

```{r, message=FALSE}
screen <- read.csv("asreview_result_sysrevemotprob.csv")
screen <- screen[screen$key %in% rayyan_res$key, ]
screen$drop_these <- FALSE
screen$drop_these[which(screen$included == 1 | screen$rayyan == "False")] <- TRUE

if(!file.exists("recs_final.csv")){
  recs_final <- unique_recs[screen$key[!screen$drop_these], ]
  write.csv(recs_final, "recs_final.csv", row.names = FALSE)
} else {
  recs_final <- read.csv("recs_final.csv")
}
```

Starting with all `r nrow(recs)` records identified through Web of Science,
duplicates were removed based on DOI matches (n = 2) and title similarity (n = 54).
Rayyan QRCI [@Ouzzani2016] identified an additional 13 duplicates.
Papers were screened based on two main criteria:
They had to address emotion regulation or a synonymous construct, and the target population must overlap with the age range of adolescence [10-24, as defined by @steinbergAgeOpportunityLessons2014].
Preliminary screening was conducted in Rayyan.
After `r sum(screen$rayyan == "True")+sum(screen$rayyan == "False")` papers were screened (`r sum(screen$rayyan == "False")` excluded),
screening continued in the free open source program ASReview [@van_de_schoot_rens_2020_3828293].
This program uses machine learning to screen articles; the algorithm used in the present study was "naive Bayes".
An additional `r sum(table(screen$included, screen$rayyan)[, 1])` papers were screened (`r sum(screen$included == 1, na.rm = TRUE)-sum(screen$rayyan == "False")` excluded),
until among the most recently screened 100 papers only 6 were excluded.
In total, `r nrow(recs_final)` papers were deemed suitable for analysis.

```{r prismachart, eval = FALSE, results='asis', fig.cap="Record screening flowchart"}
if(run_everything){
  library(tidySEM)

  lo <- matrix(c("start", "dedup", "rayyan", "asreview", "included",
                 "", "dedup2", "excluded", "", ""), nrow = 5, ncol = 2)

  nodes <- data.frame(
    name = c("start", "dedup", "dedup2", "rayyan", "asreview", "included", "excluded"), 
    label = c(
      paste0("Records identified through\nWeb of Science (n = ", nrow(recs), ")"), 
      paste0("After duplicates removed\n(n = ", nrow(unique_recs), ")"), 
      paste0("Duplicates\n(n = ", nrow(recs)-nrow(screen), ")"), 
      paste0("Articles screened for eligibility\nusing Rayyan (n = ", nrow(screen), ")"), 
      paste0("Articles screened for eligibility\nusing ASReview (n = ", nrow(screen), ")"), 
      paste0("Studies included in\nqualitative synthesis (n = ", nrow(recs_final), ")"),
      paste0("Excluded studies (n = ", nrow(screen)-nrow(recs_final), ")"))
  )

  edges <- data.frame(
    from = c("start", "dedup", "rayyan", "asreview",
             "start", "rayyan", "rayyan", "asreview"),
    to = c("dedup", "rayyan", "asreview", "included",
           "dedup2", "dedup2", "excluded", "excluded"),
    label = c("",   "", paste0("n = ", sum(screen$rayyan == "True")), paste0("n = ", sum(screen$included == 0, na.rm = TRUE)-sum(screen$rayyan == "True")),
              paste0("n = ", sum(dups)), 
              paste0("n = ", rayyan_dups),
              paste0("n = ", sum(screen$rayyan == "False")),
              paste0("n = ", sum(screen$included == 1, na.rm = TRUE)-sum(screen$rayyan == "False"))
    )
  )

  p <- graph_sem(layout = lo, nodes = nodes, edges = edges, angle = 1, rect_height = 1.2, rect_width = 1.3, text_size = 4)
  ggsave("prismachart.png", p, device = "png", width = 7, height = 4)
  ggsave("prismachart.svg", p, device = "svg", width = 7, height = 4)
}

knitr::include_graphics("prismachart.png")
```

<!--
## Corpus description

{r, eval = FALSE}
# Plot papers per year
df_plot <- as.data.frame(table(recs_final$PY), stringsAsFactors = FALSE)
df_plot$Year = as.numeric(df_plot$Var1)
ggplot(df_plot, aes(x= Year, y= Freq))+geom_point() + geom_path(group =1)+
  theme_bw() + scale_y_log10()
-->

## Analysis 1: Author keywords

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(stringr)
library(udpipe)
library(igraph)
library(wordcloud)
library(Matrix)
library(ggplot2)
library(yaml)
source("word_functions.R")
source("circle2.R")

#run_everything = FALSE
study1details <- read_yaml("study1_details.yml")
dict <- read_yaml("yaml_dict.txt")
## Look at POS tags?
if(run_everything){
  recs <- read.csv("recs_final.csv")
  recs <- as.data.table(recs)
  if(!is.data.table(recs)){
    browser()
  }
  recs[, "doc" := 1:nrow(recs)]
  study1details <- list(dim_recs = dim(recs))

  # Extract individual words
  df <- lapply(recs$DE, function(x){strsplit(x, split = "; ")[[1]]})
  df <- merge_df(recs, df, "word")

  df[, word := tolower(word)]

  # Clean
  df <- na.omit(df, cols = "word")
  number_docs_words <- c(docs = length(unique(df$doc)), words = length(unique(df$word)))
  yaml::write_yaml(number_docs_words, "study1_number_docs_words.txt")
  # Exclude words
  exclude_terms <- readLines("exclude_terms.txt")
  exclude_these <- unique(unlist(lapply(exclude_terms, grep, x = df$word)))

  df <- df[!exclude_these, ]

  # Categorize words
  res_cat <- cat_words(df$word, dict, handle_dups = "all")
  # Check coding issues
  #res_cat$dup
  #head(res_cat$unmatched)
  df <- merge_df(df, res_cat$words, "word_coded")
  saveRDS(df, "study1_df.RData")
} else {
  df <- readRDS("study1_df.RData")
  number_docs_words <- yaml::read_yaml("study1_number_docs_words.txt")
}

# Frequency of word by doc
  nounbydoc <- df[, list(freq = .N), by = list(doc_id = doc, term = word_coded)]
  # Set frequency to 1; we're not interpreting word frequency, only occurrence
  nounbydoc$freq <- 1

  dtm <- udpipe::document_term_matrix(document_term_frequencies(nounbydoc))
  term_freq <- table(colSums(dtm))
  # Continue plotting word frequency ----------------------------------------

  set.seed(5348)
  dtm_top <- dtm[, select_words(dtm, .975)]
  dtm_top <- dtm_top[rowSums(dtm_top) > 0, ]

if(run_everything){

  term_freqtop <- table(colSums(dtm_top))
  term_freq <- as.data.frame.table(term_freq)
  term_freq$pruned <- term_freq$Var1 %in% names(term_freqtop)
  write_yaml(term_freq, "study1_term_freq_dist.yml")
  write_yaml(dim(dtm_top), "study1_dtm_top.yml")

  # Wordcloud ---------------------------------------------------------------

  ## Word frequencies
  topterms <- colSums(dtm_top)
  baseline <- readRDS("baseline.RData")

  word_freq <- data.frame(Word = names(topterms), Frequency = topterms, row.names = NULL)
  write.csv(word_freq, "study1_word_freq.csv", row.names = FALSE)
  df_plot <- word_freq
  df_plot <- df_plot[order(df_plot$Frequency, decreasing = TRUE), ]
  categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
  df_plot$cat <- categ$category[match(df_plot$Word, categ$name)]
  df_plot$baseline <- as.character(df_plot$Word %in% baseline)
  df_plot$faded <- df_plot$Word %in% baseline
  # Tag words that are not in the cooccurrence graph
  in_graph <- row.names(read.csv("s1_cooc.csv", row.names = 1))
  notingraph <- !df_plot$Word %in% in_graph
  df_plot$notingraph <- notingraph
  italic_labels <- as.character(df_plot$Word)
  italic_labels[notingraph] <- sapply(italic_labels[notingraph], function(x){
    parse(text = paste0("italic('", x, "')"))
  })
  # Prettify words
  df_plot$Word <- pretty_words(df_plot$Word)

  df_plot$Word <- ordered(df_plot$Word, levels = df_plot$Word[order(df_plot$Frequency)])

  cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "forestgreen")
  df_plot$cat <- ordered(df_plot$cat, levels = c("Outcome", "Indicator", "Cause", "Protective"))

  write_yaml(df_plot$Word, "s1_words.yml")

  p <- ggplot(df_plot, aes(y = Word, x = Frequency)) +
    geom_segment(aes(x = 0, xend = Frequency,
                     y = Word, yend = Word, linetype = notingraph), colour = "grey50"
    ) + 
    geom_vline(xintercept = 0, colour = "grey50", linetype = 1) + xlab("Word frequency") +
    geom_point(data = df_plot[df_plot$faded, ], aes(colour = cat, shape = cat), fill = "white", size = 1.5) +
    geom_point(data = df_plot[!df_plot$faded, ], aes(colour = cat, fill = cat, shape = cat), size = 1.5) +
    scale_colour_manual(values = c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "forestgreen"))+ #, guide = NULL
    scale_shape_manual(values = scale_shapes)+#, guide = NULL
    scale_fill_manual(values = c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "forestgreen")) +
    scale_x_log10() +
    #scale_y_discrete(labels = italic_labels) +
    scale_linetype_manual(values = c("TRUE" = 3, "FALSE" = 1), guide = NULL) +
    theme_bw() + theme(panel.grid.major.x = element_blank(),
                       panel.grid.minor.x = element_blank(), axis.title.y = element_blank(),
                       legend.position = c(.70,.125),
                       legend.title = element_blank(),
                       axis.text.y = element_text(hjust=0, vjust = 0, size = 6))

  saveRDS(p, "s1_varimp.RData")  
  svg("s1_varimp.svg", width = 7/2.54, height = 14/2.54)
  eval(p)
  dev.off()

  ggsave("s1_varimp.png", p, device = "png", width = 7, height = 14, units = "cm")

  p1 <- p + theme(legend.position = "none")

  df_plot$Frequency <- sqrt(df_plot$Frequency)
  ## Visualise them with wordclouds
  p <- quote({
    set.seed(46)
    wordcloud(words = df_plot$Word, freq = df_plot$Frequency, scale = c(2,.4), max.words = 150, rot.per = 0,  random.order = FALSE, colors = brewer.pal(8, "Dark2"))
  })
  svg("study1_wordcloud.svg")
  eval(p)
  dev.off()

  png("study1_wordcloud.png")
  eval(p)
  dev.off()
}

# Co-occurrence -----------------------------------------------------------
set.seed(52)
cooc <- create_cooc(dtm_top)
tmp <- cooc > 0
total_cooc <- sum(tmp[lower.tri(tmp)])
cooc <- select_cooc(create_cooc(dtm_top), q = .975)
tmp <- cooc > 0
total_cooc <- c(total_cooc, sum(tmp[lower.tri(tmp)]))

if(run_everything){
  write.csv(as.matrix(cooc), "s1_cooc.csv")
  df_plot <- as_cooccurrence(cooc)
  df_plot <- df_plot[!df_plot$term1 == df_plot$term2, ]
  df_plot <- df_plot[order(df_plot$cooc, decreasing = TRUE), ]

  df_plot$id <- apply(df_plot[, c("term1", "term2")], 1, function(x)paste0(sort(x), collapse = ""))
  df_plot <- df_plot[!duplicated(df_plot$id), ]

  # Write study details -----------------------------------------------------

  write_yaml(study1details, "study1_details.yml")

  # Create network ----------------------------------------------------------

  edg <- df_plot
  edg$width = edg$cooc

  vert <- data.frame(name = names(topterms), label = pretty_words(names(topterms)), size = topterms)
  vert <- vert[vert$name %in% unique(c(edg$term1, edg$term2)), ]

  categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
  if(any(!vert$name %in% categ$name)){
    write.table(vert$name[!vert$name %in% categ$name], "clipboard", sep = "\n", row.names = FALSE, col.names= FALSE)
    stop("Please re-categorize missing vertices.")
  } 
  vert$Category <- categ$category[match(vert$name, categ$name)]
  vert$faded <- vert$name %in% baseline

  cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "olivedrab2")
  vert$color <- cat_cols[vert$Category]
  vert$shape <- cat_shapes[vert$Category]
  vert$frame.color <- cat_cols[vert$Category]
  vert$color[vert$faded] <- "#FFFFFF"

  vert$size <- scales::rescale(log(vert$size), c(4, 12))
  g <- graph_from_data_frame(edg, vertices = vert,
                             directed = FALSE)

  # edge thickness
  E(g)$width <- scales::rescale(sqrt(E(g)$width), to = c(2, 8))

  dysreg_vertex = which(names(V(g)) == "dysregulation")

  edge.start <- ends(g, es=E(g), names = FALSE)[,1]
  edge.end <- ends(g, es=E(g), names = FALSE)[,2]
  E(g)$lty <- c(1, 5)[(!(edge.start == dysreg_vertex|edge.end == dysreg_vertex))+1]

  # Layout
  set.seed(6) #4
  l1 <- l <- layout_with_fr(g)
  l1[,1] <- -1*l1[,1]
  set.seed(64)
  l2 <- layout_in_circle(g, order = shifter(V(g), -3))

  p <- quote({
    # Set margins to 0
    par(mar=c(0,0,0,0),
        mfrow=c(1,2))
    plot(g, edge.curved = 0, layout=l1,
         vertex.label.family = "sans",
         vertex.label.cex = 0.5,
         #vertex.shape = "circle2",
         #vertex.frame.color = 'gray40',
         vertex.label.color = 'black',      # Color of node names
         vertex.label.font = 1,         # Font of node names
         vertex.frame.width = 2
    )
    legend(x=-1.1, y=1.1, names(cat_cols), pch=legend_shapes, col="#777777", pt.bg=cat_cols, pt.cex=1.5, cex=.7, bty="n", ncol=1)
    plot(g, edge.curved = 0, layout=l2,
         vertex.label.family = "sans",
         vertex.label.cex = 0.5,
         #vertex.shape = "circle2",
         #vertex.frame.color = 'gray40',
         vertex.label.color = 'black',      # Color of node names
         vertex.label.font = 1,         # Font of node names
         vertex.frame.width = 2
    )
  })

  # Save files
  saveRDS(g, "s1_g1.RData")
  saveRDS(l1, "s1_l1.RData")

  png("study1_network1.png", width = 960)
  eval(p)
  dev.off()

  svg("study1_network1.svg", width = 14)
  eval(p)
  dev.off()
}
categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
word_freq <- read.csv("study1_word_freq.csv", stringsAsFactors = FALSE)
word_graph <- read.csv("s1_cooc.csv", row.names = 1)
notingraph <- word_freq$Word[!word_freq$Word %in% row.names(word_graph)]
notingraph <- categ[categ$name %in% notingraph, ]
cats <- unique(notingraph$category)
notingraph <- lapply(cats, function(x){ 
  out <- notingraph$name[notingraph$category == x]
  out <- pretty_words(out)
  paste0(paste0("*", out[-length(out)], "*", collapse = ", "), ", and *", tail(out, 1), "*")
  })
names(notingraph) <- cats

lda_dims <- read_yaml("Study1_lda_dims.txt")
dtm_top <- read_yaml("study1_dtm_top.yml")
term_freq_dist <- read_yaml("study1_term_freq_dist.yml")
```

The corpus for this first analysis consisted of author-provided keywords.
Keywords were extracted by document,
and an exclusion filter of methodological terms and similar non-substantive words was applied.
The resulting corpus consisted of `r number_docs_words[1]` documents with `r number_docs_words[2]` unique terms.
One important step in reviewing the literature is to examine heterogeneity of the corpus;
to determine whether there is, for example, a clear divide between psychiatric and developmental texts.
To this end, topic modeling using latent dirichlet allocation was conducted [@bleiLatentDirichletAllocation2003].
However, as no subcorpora could be identified (see [online supplement](https://github.com/cjvanlissa/veni_sysrev/blob/master/topic_models.md)), the sample was analyzed in its entirety.

### Identifying common terms

```{r message = FALSE, echo = FALSE, warning = FALSE, results="hide"}
p <- quote(cowplot::plot_grid(readRDS("s1_varimp.RData")+theme(legend.position = "none")+xlab("Word frequency (Analysis 1)"), readRDS("s2_varimp.RData")+xlab("Word frequency (Analysis 2)")))

png("varimps.png", height = 150, width = 200, units = "mm", res= 300)
eval(p)
dev.off()
```

To identify what phenomena are represented in this corpus,
terms occurring in the largest number of texts were analyzed.
To classify closely related terms, a dictionary was used that described `r length(dict)` terms using `r length(unlist(dict))` regular expression queries.
This dictionary and the classification function are available for reuse.
After dictionary encoding, the remaining number of unique terms was `r length(unique(df$word_coded))`.
To reduce the number of terms to a more manageable set,
word frequency was modeled with a negative binomial distribution.
Terms exceeding the $97.5^{th}$ percentile were retained.
This corresponded to terms occurring in at least `r min(as.numeric(term_freq_dist[[1]])[term_freq_dist$pruned])` documents.
This is a subjective criterion,
but compared to the common practices of either retaining a fixed number of terms or pruning terms below a fixed frequency [see @benoitQuantedaPackageQuantitative2018],
it has the advantage of being responsive to the empirical distribution of term frequencies.
Note that the vast majority of pruned terms (`r term_freq_dist$Freq[1]`) occurred only once in the corpus.
Pruning resulted in `r dtm_top[2]` remaining terms, which occurred in `r dtm_top[1]` documents.
The issues covered in this body of literature are visualized in Figure \@ref(fig:varimps).

```{r varimps, warning=FALSE, message = FALSE, eval = TRUE, results='asis', out.width = "100%", fig.cap="Frequency of terms in Analysis 1 and 2. Transparent nodes are constructs represented in theory (Figure 1), dotted lines are constructs absent from the co-occurrence graph."}
knitr::include_graphics("varimps.png") 
```

### Mapping the literature

To map the literature, a term co-occurrence matrix was computed,
which represents how frequently words occurred within the same document (see Figure \@ref(fig:tmnetworks)).
In total, there were `r total_cooc[1]` co-occurrence relationships.
To aid interpretability, small coefficients were again pruned using a negative binomial distribution,
retaining co-occurrences exceeding the $97.5^{th}$ percentile.
Note that this is a subjective criterion, which corresponded to terms that co-occurred in more than `r attr(cooc, "thres")` documents.
After pruning, `r total_cooc[2]` co-occurrence relationships remained.

Following the procedure described in the introduction to classify phenomena discussed in the theoretical literature,
each of the remaining terms were categorized as either a putative cause, outcome, protective factor, or indicator of emotion regulation.
Note that this classification is based on a subjective reading of the literature,
and that the nature of associations between phenomena is likely to be more complex than presented here.
For example, some of these associations are likely to be spurious, or bidirectional [e.g., emotion regulation is known to both predict and be predicted by conflict with parents, @vanlissaCostEmpathyParentadolescent2017].
This classification is merely intended as a starting point for further reflection on the potential role of each construct.

```{r tmnetworks, warning=FALSE, eval = TRUE, results='asis', out.width = "100%", fig.cap="Map of phenomena relevant to adolescents' emotion regulation based on term co-occurrence in author keywords (a) and abstracts (b). Size of lines and nodes represents frequency. Dashed lines represent links not involving emotion regulation. Transparent nodes indicate constructs also represented in the theoretical review."}

knitr::include_graphics("tmnetworks.png") 
```
```{r study1network, eval = FALSE, warning=FALSE, out.width = "100%", fig.cap="Study 1 term co-occurrence. Two layouts are presented to aid interpretation. Size of lines and nodes represents frequency. Dashed lines represent links not involving dysregulation. Transparent nodes indicate constructs also represented in the theory (cf Figure 1)."}
#knitr::include_graphics("study1_network1.svg")
library(cowplot)
fig_svg <- cowplot::ggdraw()+cowplot::draw_image("study1_network1.svg")
plot(fig_svg)
```

## Analysis 2: Abstract text mining

```{r}
library(stringr)
library(udpipe)
library(igraph)
library(wordcloud)
library(Matrix)
library(yaml)
library(Rmpfr)
library(topicmodels)
library(udpipe)
library(slam)
library(tidytext)
library(ggplot2)
library(textrank)
source("word_functions.R")
source("circle2.R")

if(isTRUE(tools::md5sum("study2_df.RData") == "21abb2622dfd94d8bd3473ed56697f9e")){
  df <- readRDS("study2_df.RData")
} else {
  stop("Note: Study 2 file is out of date. Please run this code manually.")
  recs <- data.table(read.csv("recs_final.csv"))
  if(!is.data.table(recs)){
    browser()
  }
  recs[, "doc" := 1:nrow(recs)]

  recs$AB <- tolower(recs$AB)
  if(!file.exists("english-ewt-ud-2.4-190531.udpipe")) {
    ud_model <- udpipe_download_model(language = "english")
  } else {
    ud_model <- udpipe_load_model("english-ewt-ud-2.4-190531.udpipe")
    ud_model <- udpipe_load_model(ud_model$file)
  }
  udp_res <- udpipe_annotate(ud_model, x = recs$AB, doc_id = recs$doc)
  df <- as.data.table(udp_res)
  saveRDS(df, "study2_df.RData")
}

# Keyword extraction ------------------------------------------------------

# Exclude words
if(!run_everything){
  df_kw <- readRDS("study2_df_kw.RData")
} else {
  df_kw <- df[upos %in% c("NOUN", "ADJ"), ]
  df_kw <- df_kw[grepl("^[a-zA-Z].", df_kw$lemma), ]
  exclude_terms <- readLines("exclude_terms.txt")
  exclude_these <- unique(unlist(lapply(exclude_terms, grep, x = df_kw$lemma)))
  df_kw <- df_kw[-exclude_these, ]
  saveRDS(df_kw, "study2_df_kw.RData")
}

# No numeric values
# all(is.na(as.numeric(df_kw$lemma)))
#df_kw$lemma[nchar(df_kw$lemma) == 3]

if(!run_everything){
  kw_tr <- readRDS("study2_textrank.RData")
} else {
  kw_tr <- textrank_keywords(x = df_kw$lemma[df_kw$upos %in% c("NOUN", "ADJ")], 
                             ngram_max = 3, sep = " ")
  saveRDS(kw_tr, "study2_textrank.RData")
}

if(!run_everything){
  df_analyze <- data.table(readRDS("study2_df_analyze.RData"))
} else {
  # Merge back with original data
  df_kw$keyword <- txt_recode_ngram(df_kw$lemma, compound = kw_tr$keywords$keyword, ngram = kw_tr$keywords$ngram, sep = " ")
  df_kw$keyword[!df_kw$keyword %in% kw_tr$keywords$keyword] <- NA

  df_analyze <- df_kw[!is.na(df_kw$keyword), ]
  dict <- read_yaml("yaml_dict.txt")
  res_cat <- cat_words(df_analyze$keyword, dict, handle_dups = "all")
  # Check coding issues
  #res_cat$dup
  #head(res_cat$unmatched)
  df_analyze <- merge_df(df_analyze, res_cat$words, "word_coded")
  saveRDS(df_analyze, "study2_df_analyze.RData")
}

# Wordcloud ---------------------------------------------------------------
# Frequency of word by doc
if(!is.data.table(df_analyze)) stop("Not a data.table")
nounbydoc <- df_analyze[, list(freq = .N), by = list(doc_id = doc_id, term = word_coded)]
number_docs_words2 <- c(docs = length(unique(nounbydoc$doc_id)), words = length(unique(nounbydoc$term)))

nounbydoc$freq <- 1
dtm <- udpipe::document_term_matrix(document_term_frequencies(nounbydoc))

topterms <- colSums(dtm)
term_freq <- table(topterms)
topterms <- sort(topterms, decreasing = TRUE)

# Select most common terms ------------------------------------------------
set.seed(720)
dtm_top <- dtm[, select_words(dtm, .975)]
dtm_top <- dtm_top[rowSums(dtm_top) > 0, ]

term_freqtop <- table(colSums(dtm_top))
term_freq <- as.data.frame.table(term_freq)
term_freq$pruned <- term_freq[[1]] %in% names(term_freqtop)
write_yaml(term_freq, "study2_term_freq_dist.yml")

if(run_everything){

  # Wordcloud ---------------------------------------------------------------

  ## Word frequencies
  topterms <- colSums(dtm_top)
  topterms <- sort(topterms, decreasing = TRUE)
  word_freq <- data.frame(Word = names(topterms), Frequency = topterms, row.names = NULL)
  write.csv(word_freq, "study2_word_freq.csv", row.names = FALSE)
  df_plot <- word_freq
  df_plot$Word <- pretty_words(df_plot$Word)
  df_plot$Frequency <- sqrt(df_plot$Frequency)
  ## Visualise them with wordclouds
  p <- quote({
    set.seed(46)
    wordcloud(words = df_plot$Word, freq = df_plot$Frequency, scale = c(4,.4), max.words = 150, rot.per = 0,  random.order = FALSE, colors = brewer.pal(8, "Dark2"))
  })

  svg("study2_wordcloud.svg")
  eval(p)
  dev.off()

  png("study2_wordcloud.png")
  eval(p)
  dev.off()
}

# Feature importance ------------------------------------------------------
topterms <- colSums(dtm_top)
if(run_everything){
  baseline <- readRDS("baseline.RData")

  word_freq <- data.frame(Word = names(topterms), Frequency = topterms, row.names = NULL)
  df_plot <- word_freq
  df_plot <- df_plot[order(df_plot$Frequency, decreasing = TRUE), ]
  categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
  df_plot$cat <- categ$category[match(df_plot$Word, categ$name)]
  df_plot$faded <- df_plot$Word %in% baseline
  # Tag words that are not in the cooccurrence graph
  in_graph <- row.names(read.csv("s2_cooc.csv", row.names = 1))
  notingraph <- !df_plot$Word %in% in_graph
  df_plot$notingraph <- notingraph
  italic_labels <- as.character(df_plot$Word)
  italic_labels[notingraph] <- sapply(italic_labels[notingraph], function(x){
    parse(text = paste0("italic('", x, "')"))
  })
  df_plot$Word <- pretty_words(df_plot$Word)

  df_plot$Word <- ordered(df_plot$Word, levels = df_plot$Word[order(df_plot$Frequency)])

  cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "forestgreen")
  df_plot$cat <- ordered(df_plot$cat, levels = c("Outcome", "Indicator", "Cause", "Protective"))

  write_yaml(df_plot$Word, "s2_words.yml")

  p <- ggplot(df_plot, aes(y = Word, x = Frequency)) +
  geom_segment(aes(x = 0, xend = Frequency,
                   y = Word, yend = Word, linetype = notingraph), colour = "grey50"
  ) + 
  geom_vline(xintercept = 0, colour = "grey50", linetype = 1) + xlab("Word frequency") +
  geom_point(data = df_plot[df_plot$faded, ], aes(colour = cat, shape = cat), fill = "white", size = 1.5) +
  geom_point(data = df_plot[!df_plot$faded, ], aes(colour = cat, fill = cat, shape = cat), size = 1.5) +
  scale_colour_manual(values = c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "forestgreen"), guide = NULL)+#
  scale_shape_manual(values = scale_shapes, guide = NULL)+#
  scale_fill_manual(values = c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "forestgreen"), guide = NULL) +
  scale_x_log10() +
  #scale_y_discrete(labels = italic_labels) +
  scale_linetype_manual(values = c("TRUE" = 3, "FALSE" = 1), guide = NULL) +
  guides(fill = guide_legend(title = NULL,
                               override.aes = list(linetype = c(1, 1, 1,1),
                                                   shape = scale_shapes,
                                                   colour = cat_cols[names(scale_shapes)]))) + 
theme_bw() + theme(panel.grid.major.x = element_blank(),
                   panel.grid.minor.x = element_blank(), axis.title.y = element_blank(),
                   legend.position = c(.825,.1125),
                   legend.title = element_blank(),
                   axis.text.y = element_text(hjust=0, vjust = 0, size = 6))
  saveRDS(p, "s2_varimp.RData")
  svg("s2_varimp.svg", width = 7/2.54, height = 14/2.54)
  eval(p)
  dev.off()

  ggsave("s2_varimp.png", p, device = "png", width = 7, height = 14, units = "cm")

}

# Co-occurrence -----------------------------------------------------------
set.seed(5646)
cooc <- create_cooc(dtm_top)
tmp <- cooc > 0
total_cooc_2 <- sum(tmp[lower.tri(tmp)])
cooc <- select_cooc(create_cooc(dtm_top), q = .975)
tmp <- cooc > 0
total_cooc_2 <- c(total_cooc_2, sum((tmp > 0)[lower.tri(tmp)]))

if(run_everything){
  write.csv(as.matrix(cooc), "s2_cooc.csv")
  df_plot <- as_cooccurrence(cooc)
  df_plot <- df_plot[!df_plot$term1 == df_plot$term2, ]
  df_plot <- df_plot[order(df_plot$cooc, decreasing = TRUE), ]

  df_plot$id <- apply(df_plot[, c("term1", "term2")], 1, function(x)paste0(sort(x), collapse = ""))
  df_plot <- df_plot[!duplicated(df_plot$id), ]

  # Create network ----------------------------------------------------------

  edg <- df_plot
  edg$width = edg$cooc

  vert <- data.frame(name = names(topterms), label = pretty_words(names(topterms)), size = topterms)
  vert <- vert[vert$name %in% unique(c(edg$term1, edg$term2)), ]

  categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
  if(any(!vert$name %in% categ$name)){
    write.table(vert$name[!vert$name %in% categ$name], "clipboard", sep = "\n", row.names = FALSE, col.names= FALSE)
    stop("Please re-categorize missing vertices.")
  } 
  vert$Category <- categ$category[match(vert$name, categ$name)]
  vert$faded <- vert$name %in% baseline

  cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "olivedrab2")
  vert$color <- cat_cols[vert$Category]
  vert$shape <- cat_shapes[vert$Category]
  vert$frame.color <- cat_cols[vert$Category]

  vert$color[vert$faded] <- "#FFFFFF"

  vert$size <- scales::rescale(log(vert$size), c(4, 12))
  g <- graph_from_data_frame(edg, vertices = vert,
                             directed = FALSE)

  # edge thickness
  E(g)$width <- scales::rescale(sqrt(E(g)$width), to = c(.2, 8))
  dysreg_vertex = which(names(V(g)) == "dysregulation")

  edge.start <- ends(g, es=E(g), names = FALSE)[,1]
  edge.end <- ends(g, es=E(g), names = FALSE)[,2]
  # Color edges based on origin:
  #E(g)$color <- V(g)$color[edge.start]
  E(g)$lty <- c(1, 5)[(!(edge.start == dysreg_vertex|edge.end == dysreg_vertex))+1]

  set.seed(5) #4 #2 #3
  l1 <- l <- layout_with_fr(g)
  set.seed(3) #64
  l2 <- layout_in_circle(g, order = shifter(V(g), -1))

  p <- quote({
    # Set margins to 0
    par(mar=c(0,0,0,0),
        mfrow=c(1,2))
    plot(g, edge.curved = 0, layout=l1,
         vertex.label.family = "sans",
         vertex.label.cex = 0.5,
         #vertex.shape = "circle2",
         #vertex.frame.color = 'gray40',
         vertex.label.color = 'black',      # Color of node names
         vertex.label.font = 1,         # Font of node names
         vertex.frame.width = 2
    )
    text(-1.1, y = -1, labels = "a")
    legend(x=-1.1, y=1.1, names(cat_cols), pch=legend_shapes, col="#777777", pt.bg=cat_cols, pt.cex=1.4, cex=.7, bty="n", ncol=1)
    plot(g, edge.curved = 0, layout=l2,
         vertex.label.family = "sans",
         vertex.label.cex = 0.5,
         #vertex.shape = "circle2",
         #vertex.frame.color = 'gray40',
         vertex.label.color = 'black',      # Color of node names
         vertex.label.font = 1,         # Font of node names
         vertex.frame.width = 2
    )
    text(-1.1, y = -1, labels = "b")
  })

  # Save files
  saveRDS(g, "s2_g1.RData")
  saveRDS(l1, "s2_l1.RData")
  png("study2_network1.png", width = 960)
  eval(p)
  dev.off()

  svg("study2_network1.svg", width = 14)
  eval(p)
  dev.off()
}

word_freq <- read.csv("study2_word_freq.csv", stringsAsFactors = FALSE)
word_graph <- read.csv("s2_cooc.csv", row.names = 1)
notingraph <- word_freq$Word[!word_freq$Word %in% row.names(word_graph)]
notingraph <- categ[categ$name %in% notingraph, ]
cats <- unique(notingraph$category)
notingraph <- lapply(cats, function(x){ 
  out <- notingraph$name[notingraph$category == x]
  out <- pretty_words(out)
  paste0(paste0("*", out[-length(out)], "*", collapse = ", "), ", and *", tail(out, 1), "*")
})
names(notingraph) <- cats
lda_dims_2 <- read_yaml("Study2_lda_dims.txt")
term_freq <- read_yaml("study2_term_freq_dist.yml")

```
```{r mergenetworks}
cat_cols <- c(Outcome = "gray50", Indicator = "tomato", Cause = "gold", Protective = "olivedrab2")
bl1 <- readRDS("bl_l1.RData")
bl2 <- readRDS("bl_l2.RData")
bg1 <- readRDS("bl_g1.RData")
bg2 <- readRDS("bl_g2.RData")
s1l <- readRDS("s1_l1.RData")
s2l <- readRDS("s2_l1.RData")
s1g <- readRDS("s1_g1.RData")
s2g <- readRDS("s2_g1.RData")
p <- quote({
  # Set margins to 0
  par(mar=c(0,0,0,0),
      mfrow=c(2,2))
  plot(bg1, edge.curved = 0,
       layout=bl1,
       vertex.label.family = "sans",
       vertex.label.cex = 0.5,
       #vertex.shape = "circle2",
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "a")
  legend(x=-1.1, y=1.1, names(cat_cols), pch=legend_shapes, col=cat_cols, pt.bg=cat_cols, pt.cex=1.4, cex=.7, bty="n", ncol=1)

  plot(bg2, edge.curved = 0,
       layout=bl2,
       vertex.label.family = "sans",
       vertex.label.cex = 0.5,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "b")
  plot(s1g, edge.curved = 0,
       layout=s1l,
       vertex.label.family = "sans",
       vertex.label.cex = 0.5,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "c")
  plot(s2g, edge.curved = 0,
       layout=s2l,
       vertex.label.family = "sans",
       vertex.label.cex = 0.5,
       #vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "d")
})

# Save files
# png("networks.png", width = 960)
# eval(p)
# dev.off()
png("networks.png", height = 200, width = 200, units = "mm", res= 300)
eval(p)
dev.off()

svg("networks.svg", width = 14)
eval(p)
dev.off()

p <- quote({
  # Set margins to 0
  par(mar=c(0.5,0.5,0.5,0.5),
      mfrow=c(1,2))
  plot(s1g, edge.curved = 0,
       layout=s1l,
       vertex.label.family = "sans",
       vertex.label.cex = 0.5,
       ##vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "a")
  legend(x=-1.2, y=1.1, names(cat_cols), pch=legend_shapes, col=cat_cols, pt.bg=cat_cols, pt.cex=1.4, cex=.6, bty="n", ncol=1)

  plot(s2g, edge.curved = 0,
       layout=s2l,
       vertex.label.family = "sans",
       vertex.label.cex = 0.5,
       ##vertex.shape = "circle2",
       #vertex.frame.color = 'gray40',
       vertex.label.color = 'black',      # Color of node names
       vertex.label.font = 1,         # Font of node names
       vertex.frame.width = 2
  )
  text(-1.1, y = -1, labels = "b")

})

png("tmnetworks.png", height = 100, width = 200, units = "mm", res= 300)
eval(p)
dev.off()

svg("tmnetworks.svg", width = 14)
eval(p)
dev.off()
```
The corpus for this second analysis consisted of the abstracts of the selected articles.
Keywords, as examined in Analysis 1, convey high-quality information because they are carefully chosen by authors to capture the essence of a study.
However, as authors are
typically limited to 5 keywords, some nuance may be lost.
Abstracts, by contrast, offer greater freedom of expression,
but present a greater challenge when it
comes to extracting relevant information.
It has been shown that retaining nouns and adjectives generally helps derive more interpretable text mining models [@martinMoreEfficientTopic2015].
For example, nouns help capture terms like "emotion", and adjectives help capture the "mental" in "mental health".
The natural language processing technique "part-of-speech tagging" (POS-tagging) was used to identify each word's grammatical function within the sentence context.
Finally, stemming was used to reduce the retained terms to their root form;
a common procedure to ensure that terms are correctly classified regardless of their use in a sentence.
As in Analysis 1, heterogeneity was explored using latent dirichlet allocation - but no subcorpora were identified ([see online supplement](https://github.com/cjvanlissa/veni_sysrev/blob/master/topic_models.md)).

### Feature engineering

When conducting text mining analysis on unstructured text data (as opposed to author key words), focusing on individual words out of context can reduce interpretability.
For instance, the central construct of this review, "emotion regulation", is already a bigram.
To obtain more meaningful units of analysis,
the `textrank` algorithm [@wijffelsTextrankSummarizeText2019] was used to identify $n$-grams (with $n \leq 3$).
This is sufficient to capture trigrams, like "parent-child conflict",
but would not capture quadragrams like "parent-child conflict resolution".
The resulting $n$-grams were merged with the original data.

### Identifying common terms

After applying the dictionary and exclusion filter as explained in Analysis 1,
the resulting corpus consisted of `r number_docs_words2[1]` documents with `r number_docs_words2[2]` unique words.
Again, the $97.5^{th}$ percentile of a negative binomial distribution was used
as a subjective threshold for pruning the least common terms,
which corresponded to terms occurring in more than `r max(as.numeric(term_freq[[1]][!term_freq$pruned]))` documents.
The identified important keywords are displayed in Figure \@ref(fig:varimps).

### Mapping the literature

A term co-occurrence matrix was constructed following the procedure described in Analysis 1.
In total, there were `r total_cooc_2[1]` co-occurrence relationships.
Small coefficients below the $97.5^{th}$ percentile of a negative binomial distribution were again pruned.
This is a subjective criterion, which corresponded to terms that co-occurred in more than `r attr(cooc, "thres")` documents.
After pruning, `r total_cooc_2[2]` co-occurrence relationships remained.
Figure \@ref(fig:tmnetworks) displays the resulting co-occurrence matrix as a force directed graph.

```{r study2network, eval = FALSE, out.width = "100%", fig.cap="Study 2 term co-occurrence. Two layouts are presented to aid interpretation. Size of lines and nodes represents frequency. Dashed lines represent links not involving dysregulation. Transparent nodes indicate constructs also represented in the theory (cf Figure 1)."}
#knitr::include_graphics("study2_network1.svg")
library(cowplot)
fig_svg <- cowplot::ggdraw()+cowplot::draw_image("study2_network1.svg")
plot(fig_svg)
```

# Results

In the analysis of author keywords (Analysis 1),
emotion regulation and associated mental health-related outcomes were foremost among the common terms in the corpus (Figure \@ref(fig:varimps)). 
Other frequent terms reflect important themes discussed in the theoretical review of the literature;
for instance, the terms *neural*, *parenting*, and *stress* correspond to themes discussed by  @coe-odessEmergentEmotionsAdolescence2019:
Neurocognitive development, the role of the parents, and adolescents' increased stress response.
Importantly, the most common terms also include several concepts not featured prominently in the theoretical review.
For example, *ADHD/CD* [cf. @braetEmotionRegulationChildren2014], *substance* use [cf. @pierrehumbertStrategiesEmotionRegulation2002], and *minority status* [cf. @myersEthnicitySocioeconomicStatusrelated2009] are common in the corpus,
but featured less prominently in the theoretical review.

```{r}
s1 <- read_yaml("s1_words.yml")
s2 <- read_yaml("s2_words.yml")

#length(unique(c(s1, s2)))

oneintwo <- round(length(s1[s1 %in% s2])/length(s1)*100)
twoinone <- round(length(s2[s2 %in% s1])/length(s2)*100)

s1order <- data.frame(word = s1, order = 1:length(s1))
s2order <- data.frame(word = s2, order = 1:length(s2))
orders <- merge(s1order, s2order, by = "word", all = TRUE)
orders[is.na(orders)] <- 999
names(orders) <- c("word", "order1", "order2")
orders$diff <- orders$order1-orders$order2
orders <- orders[order(abs(orders$diff), decreasing = TRUE), ]
```

In the analysis of abstracts (Analysis 2),
emotion regulation and associated mental health-related outcomes were evidently the most common terms.
There was substantial agreement between the most common terms identified in the analysis of author keywords and abstracts.
Specifically, $`r oneintwo`\%$ of the most frequent terms identified in the author keywords
were also present in the abstracts,
and conversely, $`r twoinone`\%$ of the most frequent terms from the abstracts were present in the keywords.
There were also some differences;
for instance, the term *sex* was more frequent than in Analysis 1,
suggesting that sex differences are regularly reported in Abstracts even if they are not mentioned in the keywords.
The term *parenting*, which ranked highly in the keyword analysis, was displaced by *mothers* in the analysis of abstracts.
This reflects the prior observation that parenting is most often operationalized in terms of mothering [@pleckPaternalInvolvementRevised2004].

<!-- Co-occurrence -->
With regard to the term co-occurrence analyses, the analysis of  keywords (Analysis 1, see Figure \@ref(fig:tmnetworks)) indicated that 
emotion regulation is evidently a central construct
to which most other constructs were directly linked.
This suggests that the search successfully identified factors relevant for adolescents’ emotion regulation.
The remaining graph was notably sparse, with few interconnections between terms.

```{r, echo = FALSE}
categ <- read.csv("study1_categorization.csv", stringsAsFactors = FALSE)
connected <- cooc[which(rownames(cooc) == "dysregulation"), ]
connected <- names(connected)[connected > 0]
connected <- connected[connected %in% categ$name[categ$category == "Cause"]]
```

Compared to the keyword co-occurrence graph,
the analysis of abstracts (Analysis 2) revealed an even sparser network. 
The structure also differed, as emotion regulation and mental health-related terms appeared to form a central axis.
Many relevant terms were connected to this axis, but not directly to
emotion regulation (as was the case in Analysis 1). 
The only terms directly connected to emotion regulation were *neural*, *mothers*, *attachment* and *ptsd*.
<!--THIS IS NOT IN THE RESULTS FOR ANALYSIS 1 
Again, many prevalent terms were absent from the co-occurrence graph,
as they were not strongly related to any other terms (see Figure \@ref(fig:varimps)). -->

# Discussion

<!-- start more broadly and give readers a good sense of the issues that need to be addressed in this area -->
<!-- highlight the complexities/questions involved and GENERALLY how you addressed them (such as in one sentence at the end of the paragraph). -->

<!-- Assume, for example, that readers have skipped much of your study and jumped straight to the discussion section. In the sentences after that, present what you found and the importance of your contribution (and limitations and so forth). I apologize for this journal style edit.-->
Although a substantial body of research has investigated emotion regulation in adolescence,
this literature is somewhat fragmented across subdisciplines and lacks a unifying theoretical framework.
The first step toward constructing such a unifying framework is identifying relevant empirical phenomena in this field.
The present study set out to identify such relevant phenomena and map associations among them using a novel method: the text mining systematic review (TMSR).

The results of both analyses
reflected some of the phenomena broadly considered to be relevant in prior theoretical literature and empirical reviews - particularly phenomena related to neurodevelopment and socialization.
Other frequently occurring phenomena were mental health-related outcomes that involve emotion regulation difficulties [see @aldaoEmotionregulationStrategiesPsychopathology2010].
This finding is consistent with emotion regulation's implication in a range of mental health problems [@leeAdolescentMentalHealth2014],
and underlines the importance of research in this area.

The analyses also identified several novel themes of relevant phenomena that are underrepresented in the theoretical literature and prior systematic reviews.
One such theme pertains to developmental disorders, such as ADHD/CD and autism.
This theme recurred in both analyses, although its constituent terms were ranked more highly in the analysis of keywords as compared to abstracts.
Another theme revolved around adolescents' physical health (sic), which was also reflected in terms like sleep, sports, and disability status.
This finding indicates that there is substantial empirical work on emotion regulation and physical health, although physical health is underrepresented in theory.
As with the theme of developmental disorders, terms related to physical health ranked more highly in the keyword analysis than in the abstract analysis.
This might indicate that a broader vocabulary was used to describe physical health in the abstracts; a diverse vocabulary increases the difficulty of correctly classifying terms belonging to a theme.
External stressors were another important theme, reflected in terms like bullying, stress, PTSD, abuse, violence, life events, historic events (e.g., earthquakes, war), parenting stress, and divorce.
Based on the review of the theoretical literature, this indeed appears to be an under-theorized area.
Conceptually, external stressors most closely align with the notion of the exosystem in the bioecological model [@bronfenbrennerBioecologicalModelHuman2007], but they have rarely been discussed as such in the theoretical literature.
The impression that external stressors are an undertheorized theme is reinforced by the fact that studies linking adverse life events to adolescent emotion regulation tend to reference mostly prior empirical work, but not a theoretical framework [e.g., @stikkelbroekAdolescentDepressionNegative2016].
Finally, one emergent theme that appears to be underrepresented in existing theory is structural disadvantage.
This theme was reflected in terms like minority status and discrimination, disability status, socio-economic status, adoption status, and sexual diversity. 
The aforementioned developmental disorders are also relevant in this context, as neuroatypical individuals tend to experience social exclusion [@cappadociaBullyingExperiencesChildren2012].
Future theoretical work might address these shortcomings.

The analysis of abstracts identified three additional themes not represented in the analysis of keywords.
The first of these themes revolved around addictive behavior (e.g., substance use and device use).
The second theme pertained to identity and moral development, two central topics in adolescent research [@meeusStudyAdolescentIdentity2011] with a common root in theory [@lapsleyMoralIdentityDevelopmental2015].
This theme is also reflected by the terms values and personality. 
A final emerging theme related to sexual development, as indicated by the terms (biological) sex, puberty, sexual diversity, and romantic relationships.
Taken together, these insights illustrate how inductive reviews can identify themes that are underemphasized in existing theory and narrative reviews.

The co-occurrence analyses revealed that most phenomena were directly linked with emotion regulation in the keyword analysis,
and with a central axis of emotion regulation and its mental health-related outcomes in the abstract analysis.
The emergence of a central axis of emotion regulation and mental health-related outcomes
again suggest that these phenomena are consistently studied together.
This makes sense, given the central role of emotion regulation in the etiology of mental health problems [@leeAdolescentMentalHealth2014].

Compared to the keyword analysis, the abstract analysis yielded a sparser network,
with fewer selected terms.
This is likely an artifact of the unstructured nature of abstracts,
which introduces greater noise in the analysis.
Thus, fewer terms will exceed the detection threshold.
In line with this interpretation, exploratory analyses indicated that 99% of terms occurred only once in the abstract analysis, compared to only 78% in the keyword analysis.
Despite the sparser network, there was substantial correspondence between the terms retained in both networks.
This supports the validity of the findings,
and suggests that automatic keyword extraction from abstracts can identify relevant constructs,
and may be a suitable alternative to author-provided keywords.

Both co-occurrence networks revealed relatively few interconnections between phenomena,
and several frequently occurring terms were pruned due to a lack of connectivity.
This provides empirical supports for the prior anecdotal claim that this literature is somewhat fragmented [@riedigerEmotionRegulationAdolescence2014]. 
The absence of terms from the co-occurrence networks does not mean that these phenomena are unimportant, but rather, that they are not well-integrated in the broader literature on adolescents' emotion regulation.
Several of these unembedded phenomena represent emerging themes in the literature; as is the case for research on fathers [see @pleckPaternalInvolvementRevised2004], identity [@campbellFriendsEducationIdentity2019], friendship and social support [@wangBidirectionalEffectsExpressive2020], autonomy [@vrolijkLongitudinalLinkagesFather2020a], sexual risk [@brownDepressiveSymptomsPredictor2006], and loneliness [@spithovenItAllTheir2017].

## Implications

This results of this text mining systematic review reflected several familiar phenomena that are well-represented in existing theory.
This has implications for the validity of the method, as it suggests that text mining can indeed be used to map relevant themes in the literature.
The results also revealed relevant phenomena that
are under-emphasized in narrative reviews and theory, such as developmental disorders, physical health, external stressors, structural disadvantage, substance use, and identity-, moral-, and sexual development.
This has implication for future empirical research:
In designing a study, researchers typically rely on theoretical foundations - and under-theorized phenomena are at risk of being overlooked.
The present study offers guidance regarding additional relevant phenomena to consider as potential confounders or contributing causes.

The results further revealed that several relevant phenomena were relatively unembedded:
These scored highly on term frequency,
but were sparsely connected to other terms in the co-occurrence networks, or pruned from the graph altogether.
This finding has implications for future research, as unembedded terms indicate potential knowledge gaps.
Future research might study these phenomena in conjunction with other more well-established constructs
in order to bridge gaps in existing knowledge and bring these phenomena into the mainstream literature.

Additional implications arise from the structure of the co-occurrence graphs.
Both analyses revealed close ties between emotion regulation and mental health-related outcomes, which is consistent with the role of emotion regulation in the etiology of mental health problems [see @leeAdolescentMentalHealth2014].
This has implications for the societal relevance of research in this area.
Furthermore, both analyses revealed that most constructs were
directly tied to emotion regulation and mental health,
with few connections among constructs.
These sparse networks validate prior anecdotal claims
that the literature in this field is somewhat fragmented [@riedigerEmotionRegulationAdolescence2014]
and lacks an overarching theoretical framework [@bussTheoriesEmotionalDevelopment2019].
An important implication of this finding is that, indeed, work remains to be done to integrate this diffuse field, and to generate an overarching theory of adolescent emotion regulation.
The first step in theory construction methodology (TCM) is to identify relevant phenomena [@borsboomTheoryConstructionMethodology2020].
The present study takes that first step
by mapping the relevant phenomena in this area of research using a relatively comprehensive and objective method.<!--The relevant phenomena identified in the present study might thus serve as a starting point for future theory development.-->

An important direction for future research would be to formalize these inductive insights into a new overarching theory of adolescent emotion regulation.
This would require describing and quantifying the nature of the relationship of each phenomenon in relation to adolescent emotion regulation [@borsboomTheoryConstructionMethodology2020].
This involves abductive reasoning: the attribution of observations to causal explanatory principles [@haigExploratoryFactorAnalysis2005].
As a starting point for this effort, the co-occurrence graphs presented here could be
used as templates for a nomological network:
a proto-theoretical diagrammatic representation that describes causal relationships between
relevant phenomena [see @alaviAligningTheoryMethodology2018].

```{r countreferences, eval = FALSE}
count_refs <- function(txt){
  tmp <- paste0(txt, collapse = "\r\n")
tmp <- strsplit(tmp, "\r\n")[[1]]
tmp <- tmp[grep("References", tmp):length(tmp)]
sum(stringr::str_count(tmp, "\\(\\d{4}.?\\)\\."))
}
flz <- c(
  "C:/Users/lissa102/Zotero/storage/CVUUAUGU/Bariola et al. - 2011 - Child and adolescent emotion regulation The role .pdf",
  "C:/Users/lissa102/Zotero/storage/GR3XBSYH/Coe-Odess et al. - 2019 - Emergent Emotions in Adolescence.pdf",
  "C:/Users/lissa102/Zotero/storage/P6B5XMVB/Riediger_Klipker_2014_Emotion regulation in adolescence.pdf",
  "C:/Users/lissa102/Zotero/storage/WHFYNBBN/Stifter and Augustine - 2019 - Emotion Regulation.pdf"
)
refs <- sapply(flz, function(x){
  count_refs(pdftools::pdf_text(x))
})
flz[which.max(refs)]
max(refs)
```

## Stengths and Limitations

One strength of the present study is that it was more comprehensive than previous reviews of the literature in two respects [cf. @bariolaChildAdolescentEmotion2011; @coe-odessEmergentEmotionsAdolescence2019]:
First, in contrast to prior literature reviews, the present study was based on a systematic literature search.
Second, prior reviews relied on narrative synthesis of a comparatively smaller number of publications, at most 241 [@coe-odessEmergentEmotionsAdolescence2019].
By contrast, the present study was able to map the literature more comprehensively, using text mining analysis to synthesize all 6305 identified records.
Text mining offers unique advantages compared to narrative reviews [@littellEvidencebasedBiasedQuality2008],
because it can cover vastly greater corpora than a sentient reader,
and follows a somewhat more objective, transparent, and reproducible procedure.
Both narrative and text mining reviews are initially labor-intensive.
The key differences are that text mining systematic reviews are more scalable,
because code can be applied to corpora of arbitrary size,
and are easily updated when new literature is published.
Once written, analysis code can be repurposed for reviews in different areas of research with relatively little effort.
The present study illustrates that text mining systematic reviews extract relevant information from the published literature in a way that complements theory and narrative reviews.
Moreover, text mining systematic reviews appear to be a suitable method for identifying relevant phenomena in a particular area of research.
This can serve as the first step toward formal theory construction [@borsboomTheoryConstructionMethodology2020].
The present study was the first to develop and implement this novel approach.

It should be noted that the present study has limitations as well.
The key disadvantage of the method used is that these inductive text mining techniques
are not able to extract *meaning* from the literature in the way a sentient reader would.
This limitation is best addressed by considering the results as a starting point for further abductive reasoning and theory formation, deductive empirical research, or a narrative review focused on the emergent relevant phenomena.
A related limitation is that the text mining methods used do not capture the nature of the relationship between co-occurring terms.
Instead, terms were manually classified as potential causes, outcomes, protective factors, or indicators, based on domain knowledge.
Efforts are underway to develop unsupervised algorithms capable of distilling implied causal links from scientific abstracts [@anExtractingCausalRelations2019].
Future research might substantially advance theory formation by applying such methods to the present corpus.
<!--For example, a human reader would understand that the influence of mothers and fathers both
falls under the header of "parenting". 
Furthermore, a human reader would swiftly learn that the term "parenting" has historically been used primarily to represent maternal influences,
and has increasingly come to represent paternal influence as well. -->
Another limitation is that the present analyses were limited to keywords and abstracts.
The primary obstacles to the analysis of full-text publications were limited access to articles behind paywalls,
and the absence of a standardized Application Programming Interface (API) for the automatic retrieval of articles across publishing outlets.
<!--As the number of scientific publications on any given topic tends to show exponential increase over time, making it difficult for applied researchers to keep up.
Meta-scientific analyses of the published literature will likely become increasingly important -->
Comprehensive access to scientific publications is essential to all meta-scientific research, including narrative reviews and meta-analyses (e.g., to avoid bias in summary effect sizes).
For text mining approaches in particular, the availability of large data sets is crucial.
<!--Consequently, limited access to full-text publications is a major bottleneck.-->
One example of a data set suitable for even more complex text mining analyses is a corpus of 400k+ full-text papers on COVID-19 that was made publicly available [@wangCORD19Covid19Open2020].
Compiling such full-text data sets currently requires substantial manpower and financial resources, thus placing them out of reach for many researchers.
<!--The authors argued that the sheer volume of available literature made it difficult for individual researchers to remain informed, thus necessitating automated text mining analysis.
Although our analysis comprised a mere 6305 publications, even this is -->
<!--As the number of scientific publications tends to grow exponentially across fields [@larsenRateGrowthScientific2010],
it becomes harder for researchers to remain informed, thus necessitating meta-scientific research.-->
Increased adoption of open access publishing,
and the development of a unified API for article retrieval across publishing outlets,
would facilitate curating full-text data sets.
These changes would enable more widespread adoption and more informative application of meta-scientific (text mining) analyses.
One final limitation is that several subjective decisions were made throughout the analysis process.
This creative license is inherent to inductive studies, however [@wagenmakersCreativityVerificationCyclePsychological2018], as there are infinite possible ways to conduct exploratory analyses.
To address this limitation, subjective decisions are explicitly discussed throughout the manuscript,
and all analysis code and data are made publicly available in a fully reproducible format [based on @vanlissaWORCSWorkflowOpen2020],
so that others may explore alternative exploratory analysis strategies.

# Conclusion

<!-- Begin by (briefly) reminding the reader of the topic
this paper set out to map the factors associated with adolescents' emotion regulation, based on a systematic review of the literature, and text mining analysis of author keywords and abstracts.
-->
There has been a call for the integration of the empirical literature on adolescent emotion regulation into an overarching framework.
The first step towards this end is identifying relevant phenomena in this area of research.
<!--, the research gap, and how you addressed it.-->
This paper used a text mining systematic review to identify phenomena relevant to adolescents' emotion regulation based on term frequency,
and to map potential associations between phenomena based on term co-occurrence.
<!--Then go to what you found and its importance.  Be sure to state your findings very clearly. -->
The findings echoed several themes featured in the theoretical literature review,
such as socialization and neurocognitive factors,
which illustrates the validity of this new method for identifying relevant phenomena.
Importantly, the findings also revealed undertheorized themes,
such as developmental disorders, physical health, external stressors,
structural disadvantage, substance use, identity, moral, and
sexual development.
The findings illustrate how text mining systematic reviews may complement narrative reviews<!--,
and sensitize researchers to potential confounders and contributing causes to consider in future empirical work-->.
Several of the identified themes are particularly relevant in adolescence;
either because they first emerge in this life phase (e.g., sexual development), or because they relate to age-graded developmental challenges (e.g., identity development).
This reinforces the prior observation that general theories of emotion regulation cannot simply be generalized to adolescence,<!-- a life phase qualitatively distinct from childhood and adulthood,--> and emphasizes the importance of developing theory specific to adolescence.
<!-- The implications can focus on many aspects of the study, such as the practice implications (briefly).-->
<!--the implications also should include a focus on the understanding of adolescence/early adulthood/emerging adulthood and the contribution to that research.-->
Future theoretical work might integrate these undertheorized themes into
an overarching framework, and empirical research might consider them as
promising areas for future research, or as potential confounders and contributing causes in research
on adolescents' emotion regulation.

# Acknowledgements

## Acknowledgement

Lukas Beinhauer contributed substantially to the screening of articles.

## Conflicts of Interest

The author reports no conflict of interests.

## Preregistration

This study was wholly exploratory. The analyses were not preregistered.
A full historical record of the analyses is available at https://github.com/cjvanlissa/veni_sysrev.

## Funding

This work is supported by a NWO Veni grant (NWO grant number VI.Veni.191G.090),
awarded to the author.

<!--
Informally, much can be gained by revisiting existing relevant theory
and incorporating relevant phenomena as identified in the present analysis.-->
```{r, eval = FALSE}
# These analyses takes very long (~30m); set eval = TRUE to re-run them.
if(run_everything){
  rmarkdown::render("topic_models.Rmd")
}
```

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup